### 2021-03-25 reports/20210325Manifolds.Rmd
Summary on videos about manifolds, topology

---
title: "Manifolds, topological space, and tangent space"
# author:
# - familyname: Cheng
#   othernames: Fan
#   address: Monash University
#   email: Fan.Cheng@monash.edu
blind: true
cover: false
toc: false  
bibliography: references.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: no
    number_sections: yes
    citation_package: biblatex
    latex_engine: pdflatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, cache=FALSE, messages=FALSE, warning=FALSE, eval=TRUE, fig.pos = 'H')
# options(tinytex.verbose = TRUE)
```

```{r packages, include=FALSE, eval=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
```

# Online Tutorials

YouTube tutorials on topology, manifolds & differential geometry by [*WHYBmaths*](https://www.youtube.com/playlist?list=PLxBAVPVHJPcrNrcEBKbqC_ykiVqfxZgNl)

- Preliminary: Sets & Maps

- Topology: Introduction, Constructing Torus, Topological Spaces, Continuity, Metric Spaces, Hausdorff property

- Manifolds: Introduction, Topological Manifolds, Transition Functions, Tangent Space

- Other: Vector Fields, Differential Forms

**Manifolds** are **topological spaces** that are **locally homeomorphic** to Euclidean space.

# Topological spaces

Topology space: a set $M$ with *topology* $\{\mathcal{T}_M\}$.

## Topology

Topology is the mathematical study of the properties of a geometric object that are preserved under *continuous deformations*, such as stretching, twisting, crumpling and bending. [*Wikipedia*](https://en.wikipedia.org/wiki/Topology)

Topological invariant: *homeomorphic* in the property of the space.

Homeomorphism: a continuous function between topological spaces that has a continuous inverse function.

For example: a circle is homeomorphic to a rectangle/square by certain stretching.

Constructing torus in 2-D.

```{r , fig.align = 'center', out.width = "100%", echo = FALSE, eval=TRUE}
knitr::include_graphics("figures/manifold/IMG_1342.jpg")
```

## Topological space: $(M, \{\mathcal{T}_M\})$

Topology properties:

1) $M,\phi \in \{\mathcal{T}_M\}$;

2) For all topological element $\tau_i \in \{\mathcal{T}_M\}$, the finite intersection set is also in the topology,
$$
\bigcap_i \tau_i \in \{\mathcal{T}_M\};
$$

3) Similarly, the infinite union set is also in the topology,
$$
\bigcup_i \tau_i \in \{\mathcal{T}_M\};
$$

Two extreme cases:

- Chaotic topology: $\mathcal{T}_M = \{M,\phi\}$.

- Discrete topology: $\mathcal{T}_M = \{P(M),\phi\}$, where $P(M)$ is the power set containing all possible subset of M.

Standard topology $T_{std}$ of a set $U \subset \mathbb{R}^2$:

1) Consider constructing a ball $B_r(x_i)$;

2) For any open set $U \in T_{std}$, there exists a ball $B_r(U) \subset U$.

1)&2) give a 2-d topology space $(U, B_r(U))$. [Example]

# Continuity of maps

```{r , fig.align = 'center', out.width = "100%", echo = FALSE, eval=TRUE}
knitr::include_graphics("figures/manifold/IMG_1343.jpg")
```

# Metric Spaces

```{r , fig.align = 'center', out.width = "100%", echo = FALSE, eval=TRUE}
knitr::include_graphics("figures/manifold/IMG_1344.jpg")
```

# Manifolds

```{r , fig.align = 'center', out.width = "100%", echo = FALSE, eval=TRUE}
knitr::include_graphics("figures/manifold/IMG_1345.jpg")
knitr::include_graphics("figures/manifold/IMG_1346.jpg")
```

# Transition functions

```{r , fig.align = 'center', out.width = "100%", echo = FALSE, eval=TRUE}
knitr::include_graphics("figures/manifold/IMG_1347.jpg")
```

## Example
```{r , fig.align = 'center', out.width = "100%", echo = FALSE, eval=TRUE}
knitr::include_graphics("figures/manifold/IMG_1348.jpg")
```

# Tangent spaces

```{r , fig.align = 'center', out.width = "100%", echo = FALSE, eval=TRUE}
knitr::include_graphics("figures/manifold/IMG_1349.jpg")
```

# Others

## Vector fields

```{r , fig.align = 'center', out.width = "100%", echo = FALSE, eval=TRUE}
knitr::include_graphics("figures/manifold/IMG_1350.jpg")
```

## Differential forms

```{r , fig.align = 'center', out.width = "100%", echo = FALSE, eval=TRUE}
knitr::include_graphics("figures/manifold/IMG_1351.jpg")
```