### 2023-06-05 jcgs/kderm_jcgs.Rmd
Formatting and citations

---
title: "Distortion corrected kernel density estimate on Riemannian manifolds"
blinded: 0 # Set to 1 for blinding
authors:
  - name: Fan Cheng
    affiliation: Monash University
    thanks: Fan.Cheng@monash.edu
  - name: Rob J Hyndman
    affiliation: Monash University
  - name: Anastasios Panagiotelis
    affiliation: The University of Sydney
abstract: "Manifold learning can be used to obtain a low-dimensional representation of the underlying Riemannian manifold given the high-dimensional data. However, kernel density estimates of the low-dimensional embedding with a fixed bandwidth fail to account for the way manifold learning algorithms distort the geometry of the Riemannian manifold. We propose a novel distortion-corrected kernel density estimator (DC-KDE) for any manifold learning embedding by introducing the estimated Riemannian metric of each point to fix the distortion in the line and volume elements. The geometric information of the manifold guarantees a more accurate density estimation of the true manifold, which subsequently could be used for anomaly detection. To compare our proposed estimator with a fixed-bandwidth kernel density estimator, we run two simulations with 2-D data from the Gaussian mixture model mapped into a 3-D twin peaks shape and a 5-D semi-hypersphere mapped in a 100-D space. We demonstrate that the proposed DC-KDE could improve the density estimates given a good manifold learning embedding and has higher rank correlations with the true manifold density. A shiny app in R is also developed for various simulation scenarios. The proposed method is applied to density estimation in statistical manifolds of electricity usage with the Irish smart meter data. This demonstrates the estimator's capability to fix the distortion of the manifold geometry and a new approach to anomaly detection in high-dimensional data."
keywords:
  - manifold learning
  - variable bandwidth
  - Riemannian metric
  - geodesic distance
  - Gaussian kernels
bibliography: kderm.bib
output:
  bookdown::pdf_document2:
    template: asa_template.tex
    citation_package: biblatex
    bib_engine: biber
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    dev: "pdf"
    keep_tex: yes
    includes:
      in_header: preamble.tex
# after_body: appendix.tex
link-citations: yes
---

```{r child = "03-chap3-kderm.Rmd"}

```

\clearpage

```{r child="B-appB-kderm.Rmd"}

```

\clearpage

# References {-}