### 2024-09-11 jcgs/03-chap3-kderm.tex
minor changes

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\hypertarget{dckdeintro}{%
\section{Introduction}\label{dckdeintro}}

Multivariate kernel density estimation (KDE see @Parzen1962-gt;
@Scott2015-vl and references therein) is a fundamental tool for
exploratory data analysis and unsupervised learning. Applications of KDE
include finding hot spots of traffic networks in the GIS environment
{[}@Xie2008-eb; @Okabe2009-nb{]}, automatic detection in visual
surveillance systems {[}@Elgammal2002-cw{]}, wind power density
detection {[}@Jeon2012-ac{]} and prime prediction via Twitter messages
{[}@Gerber2014-tq{]}. For \(m\)-dimensional data where \(m\) is even of
moderate size, KDE is challenging due to the well-known curse of
dimensionality. However, the manifold hypothesis suggests that many
high-dimensional datasets belong to a \(d\)-dimensional manifold \(M\)
where \(d\ll m\). The problem of KDE on manifolds has been well studied
in @Hendriks1990-bt, @Izenman1991-cj, @Pelletier2005-vu, @Henry2013-da,
and @Berenfeld2022-fd. However, in these settings, it is assumed that
the underlying manifold is known, which is rarely true in practice. The
is only a small literature dealing with the case where the manifold is
unknown. Recently, @berenfeld2021-den, @berenfeld2022-est and
@divol2022measure demonstrated the minimax optimality of certain types
of KDE on manifolds. In contrast, our objective in this paper is to
propose a new KDE for data that lie on some unknown manifold found using
embedding techniques.

When dealing with data that lie on a manifold, it is common to use
dimension-reduction techniques. These algorithms, which include ISOMAP,
LLE, t-SNE, and UMAP among others, can be thought of as a mapping
\(\psi:M\rightarrow \mathbb{R}^d\) that take points on the manifold
\(\bm{p}\) to \(d\)-dimensional vectors \(\bm{y}\), where the dimension
\(d\) is either assumed to be known a priori or needs to be estimated
separately. Details of these algorithms are summarized in
@Cayton2005-dp, @Lee2007-wq, @Izenman2012-mx, and @Cheng2021-dh. One
could easily apply KDE to the output vectors \(\bm{y}\). However, this
would be a misleading representation of the density of points as they
lie on the manifold. To see how, consider that the density on the
manifold must be defined with respect to an appropriate volume form,
that is \[Pr(\bm{p}\in\mathcal{A})=\int_\mathcal{A} f(\bm{p}) dvol_M,\]
where \(f\) is the probability density function on the manifold,
\(\bm{p}\) is a point on the manifold, \(\mathcal{A}\) is some region on
the manifold and \(dvol_M\) is a volume form on the manifold \(M\). By a
change of variables, this would imply the following density for
\(\bm{y}\), \begin{equation}
\label{eq:changevar}
Pr(\bm{y}\in\psi(\mathcal{A}))=\int\limits_{\psi(\mathcal{A})} (f\circ\psi^{-1})(\bm{y})|\det \bm{H}(\bm{y})|^{1/2}d\bm{y},
\end{equation} where \(\psi(\mathcal{A})\) is the image of
\(\mathcal{A}\) under \(\psi\) and \(\bm{H}(\bm{y})\) is the Riemannian
metric expressed in local coordinates given by the mapping \(\psi\)
{[}@doCarmo1992-rm{]}. It can be seen that the density in terms of
\(\bm{y}\), differs from the density on the manifold due to the term
\(|\bm{H}(\bm{y})|^{1/2}\) which can be considered roughly analogous to
the Jacobian determinant. Since \(\psi\) will typically be a non-linear
function, the value of \(|\bm{H}(\bm{y})|^{1/2}\) will change across the
manifold, in a way that measures the distortion induced by the dimension
reduction algorithm. Notably, different choices of dimension reduction
algorithm, imply different \(\psi\), meaning that if KDE is applied to
the output \(\bm{y}\) coming from two different dimension reduction
algorithms, this could lead to two very different sets of conclusions
about the density of the same point.

In this paper, we propose a new KDE. Our estimator is also based on the
output of dimension reduction algorithms but accounts for the distortion
induced by the algorithm. For this reason, we refer to it as the
distortion-corrected KDE (DC-KDE). Critical to our estimator, is
obtaining an estimate of \(|\bm{H}(\bm{y})|^{1/2}\) for this purpose we
use the Learn Metric algorithm of @Perrault-Joncas2013-pq which augments
any dimension reduction algorithm with an estimate of the Riemannian
metric at each data point. One particular downstream task that we are
interested in is anomaly detection, where anomalies are identified as
points of low density. In a number of simulation studies, we demonstrate
two benefits that arise when using our new estimator. The first is that
using distortion correction leads to more accurate density estimates and
better detection of anomalies compared to the case where distortion
correction is not used. The second is that when the estimates are found
using different dimension reduction algorithms (e.g.~ISOMAP v UMAP), the
results regarding the density estimates and the detection of anomalies
are robust only when distortion correction is applied.

In terms of data visualization, our approach allows a two or
three-dimensional scatterplot of \(\bm{y}\) to be augmented by coloring
points according to the magnitude of the density estimate, where the
dimension of \(\bm{y}\), \(d\), is set as 2 or 3 in the dimension
reduction algorithms. In this way, the user is able to tell whether a
region that has a cluster of points in the output space \(\bm{y}\) is
truly a high-density region or if instead, this is an artifact of the
dimension reduction algorithm `squeezing' points together. This is
particularly pertinent in the case of t-SNE which can form clusters of
points even when no such clusters exist in the underlying data.
Furthermore, our estimator can be used for manifolds with elements that
are a variety of objects, for instance, rotational frames or images. In
our application we consider a probabilistic manifold whose elements are
densities, meaning that we estimate a density of densities. Finally, our
proposed estimator also takes the form of a KDE with variable bandwidth
meaning that it also contributes to extensive literature on this topic
{[}see Section 6.6 of @Scott2015-vl for a summary{]}.

The rest of the paper is organized as follows. In \autoref{kderm}, we
present our distortion-corrected kernel density estimator for Riemannian
manifolds. Our estimator combines insights from two existing papers, the
first is @Pelletier2005-vu who propose a KDE for data on manifolds that
are known, the second is @Perrault-Joncas2013-pq who propose the Learn
Metric algorithm which we use to estimate the distortion induced by a
dimension reduction algorithm. \autoref{simulation} is composed of two
simulations; the first deals with 2-dimensional data from a Gaussian
mixture model mapped onto a 3-dimensional `twin peaks' manifold and the
second with a 5-dimensional semi-hypersphere data embedded into a
100-dimensional space. \autoref{dckdeapplication} contains an
application to an Irish smart meter dataset that visualizes and
identifies households with an anomalous distribution of energy usage.
Conclusions and discussions are presented in \autoref{dckdeconclusion}.
Some additional background on important terms in Riemannian geometry
required to understand our proposed method is provided in
\autoref{riemgeo}.

\hypertarget{kderm}{%
\section{Distortion Corrected Kernel density estimate on Riemannian
manifolds}\label{kderm}}

In this section, we introduce our method for kernel density estimation
on manifolds that uses an embedding from a dimension reduction algorithm
while correcting for the distortion induced by this embedding. Since
some readers may be unfamiliar with the nuances of manifolds, we first
discuss kernel density estimation for data in Euclidean space, then
illustrate in \autoref{Pellet} how this generalizes to the estimator of
@Pelletier2005-vu, which can be applied when the data lie on some known
manifold. In \autoref{MetLearn}, we describe the Learn Metric algorithm
of @Perrault-Joncas2013-pq, which augments an embedding derived from a
dimension reduction algorithm with an estimate of the Riemannian metric
expressed in local coordinates. By combining elements from the work of
@Pelletier2005-vu and @Perrault-Joncas2013-pq, we derive our own novel
distortion-corrected kernel density estimate in \autoref{DCKDE}. To keep
this section as succinct as possible, we do not define concepts such as
manifolds, charts, geodesic distance, etc., but provide this information
for readers unfamiliar with differential geometry in \autoref{riemgeo}.

In the following, we denote \(M\) as the \(d\)-dimensional manifold from
which our data are sampled. Points on this manifold are denoted
\(\bm{p}\) in general, with \(\bm{p}_1,\dots,\bm{p}_n\) denoting the
observed sample. Often \(\bm{p}_i\) will be high-dimensional vectors
such that \(\bm{p}_i\in\mathbb{R}^m\) with \(m\gg d\), however, this
need not be the case. For instance, \(\bm{p}_i\) may be probability
distributions on a statistical manifold. The methods we propose for
estimating the density at each \(\bm{p}_i\) only require some sense of
distance between the `input' points, \(d(\bm{p}_i,\bm{p}_j)\), such that
we can apply dimension reduction algorithms to obtain an `output'
embedding, \(\bm{y}_1,\dots,\bm{y}_n\), where
\(\bm{y}_i\in\mathbb{R}^d\). We will denote this embedding as
\(\bm{y}_i=\psi(\bm{p}_i)\). Finally, we denote by \(\lambda\) the
Lebesgue measure of \(\mathbb{R}^d\), letting \(\|\cdot\|\) be the usual
Euclidean norm and following @Pelletier2005-vu, we make the following
assumptions about the kernel function
\(K:\mathbb{R}_+\rightarrow\mathbb{R}\), \begin{equation}
\begin{aligned}
\label{eq:kernelcondition}
& (i) \int_{\mathbb{R}^{d}} K(\|\bm{y}\|) \mathrm{d} \lambda(\bm{y})=1;
(ii) \int_{\mathbb{R}^{d}} \bm{y} K(\|\bm{y}\|) \mathrm{d} \lambda(\bm{y})=0;
(iii) \int_{\mathbb{R}^{d}}\|\bm{y}\|^{2} K(\|\bm{y}\|) \mathrm{d} \lambda(\bm{y})<\infty; \\
& (iv) \supp K=[0; 1];
(v) \sup K(\|\bm{y}\|)=K(0).
\end{aligned}
\end{equation}

Note that these conditions are different from (and in some cases
stricter than) those normally used for kernel density estimation. For
instance, condition (iv) requires the support of the kernel to be
bounded. The reasons for this will become clearer when we discuss the
manifold setting in more detail. Also, for illustration purposes, in
this section, we pay particular attention to the uniform kernel for
which \(K(z)\) equals one if \(0\leq z\leq1\) and zero otherwise. In our
empirical section, more general kernel functions can be, and are,
employed.

For data \(\bm{y}_i\in\mathbb{R}^d\) with \(i=1,\dots,N\) and assuming a
bandwidth matrix \(r\bm{I}\) where \(r\) is a global bandwidth, then the
usual kernel density estimator at a point \(\bm{y}\) is given by
\begin{equation}
\label{eq:vkde}
\hat{f}(\bm{y})=\frac{1}{N}\sum\limits_{i=1}^N \frac{1}{r^d} K\left(\frac{\|\bm{y}-\bm{y}_i\|}{r}\right).
\end{equation}

The intuition behind this estimator is very clear for a uniform kernel.
The density at a point \(\bm{y}\) is equal to the proportion of sample
points that lie within a ball of radius \(r\) centered at \(\bm{y}\),
times a term that ensures the density integrates to 1. In general, the
bandwidth matrix need not be proportional to the identity matrix.
However, the intuition remains the same, only that the ball of radius
\(r\) centered at \(\bm{y}\) is found with respect to Mahalanobis
distance rather than the usual Euclidean distance. For more on kernel
density estimation in the Euclidean case, see @Scott2015-vl and
references therein. Kernel density estimators of this form can and have
been applied directly on the output embedding \(\bm{y}\), and we will
consider this approach as a benchmark in \autoref{simulation}.

\hypertarget{Pellet}{%
\subsection{Kernel Density estimation on manifolds}\label{Pellet}}

For kernel density estimation on a known manifold, @Pelletier2005-vu
propose the following estimator, \begin{equation}
\label{eq:denriem}
\hat{f}(\bm{p}) = \frac{1}{N} \sum_{i=1}^{N} \frac{1}{r^d \theta_{\bm{p}_i}(\bm{p})} K\left(\frac{d_g(\bm{p}, \bm{p}_i)}{r}\right),
\end{equation}

where \(d_g(\bm{p}, \bm{p}_i)\) denotes the geodesic distance between
two points on the manifold \(\bm{p}\) and \(\bm{p}_i\) and
\(\theta_{\bm{p}_i}(\bm{p})\) is known as the volume density function.
The intuition behind the term
\(K\left(\frac{d_g(\bm{p}, \bm{p}_i)}{r}\right)\) is relatively clear.
For example, for a uniform kernel, the estimator at point \(\bm{p}\)
will still depend on the proportion of sample points within a ball of
radius \(r\) centered at \(\bm{p}\). However, in this case, the geodesic
distance on the manifold is used, rather than Euclidean or Mahalanobis
distance. An additional technical assumption is that \(r\) is less than
the injectivity radius of the manifold. A definition of the injectivity
radius is given by @Chavel2006-mp and also provided in the appendix. For
our purposes, it is sufficient to note that this assumption precludes
the possibility that the radius of a ball around \(\bm{p}\) is so large
that some points `fall inside' the ball more than once. For example on a
sphere, a ball with a radius greater than half the circumference of a
great circle will wrap back around the sphere. This phenomenon also
explains why the kernel function must be bounded for density estimation
on manifolds.

The inclusion of the volume density function
\(\theta_{\bm{p}_i}(\bm{p})\) is perhaps not as immediately clear,
therefore, before providing formal details, we will briefly discuss the
intuition behind the inclusion of this term. We have already highlighted
that when using a uniform kernel, the kernel density estimate at a point
\(\bm{p}\) directly depends on the proportion of sample points within a
ball of radius \(r\) around \(\bm{p}\). However, the volume of this ball
must also be taken into account. In Euclidean space with the usual
Lebesgue measure, a radius \(r\) ball will always have the same volume
regardless of its center. The same does not hold for manifolds and
therefore the volume density function must be included in the KDE to
ensure that the estimated density integrates to one.

More formally, the volume density function can be explained as follows.
Consider the exponential map around \(\bm{p}\), given by
\(exp_{\bm{p}}(\bm{v})\), mapping vectors in the tangent space,
\(\bm{v}\in T_{\bm{p}}M\), to points on the manifold, \(\bm{q}\in M\).
Loosely, \(\bm{v}\) `points' in the direction of the geodesic between
\(\bm{p}\) and \(\bm{q}\) and traveling along this geodesic at uniform
speed \(\|\bm{v}\|\) takes place in one unit of time. Now, consider a
chart \(\varphi\) mapping points in the neighborhood of \(\bm{p}\), via
the inverse of the exponential map, to these \(\bm{v}\) vectors,
expressed in some local coordinate system. The volume density function
is then the square root of the determinant of the Riemannian metric
expressed in this coordinate system. In proving that the estimator
integrates to 1, the volume density function cancels with the term
\(|\bm{H}(\bm{y})|^{1/2}\) in Equation @ref(eq:changevar), where
\(\psi\) is the logarithmic map around \(\bm{p}_i\). For more on the
volume density function, see @Gallot2004-rc and @Le\_Brigant2019-lj.

\hypertarget{MetLearn}{%
\subsection{Riemannian metric estimation}\label{MetLearn}}

To be able to apply the estimator of @Pelletier2005-vu to the case where
the manifold is not known, but where coordinates \(\bm{y}_i\) for
\(i=1,\dots,n\) are obtained from a dimension reduction algorithm,
requires an estimate of the Riemannian metric. Formally, the Riemannian
metric is a symmetric and positive definite tensor field that defines an
inner product on the tangent space \(T_{\bm{p}} M\) for every point
\(\bm{p} \in M\). Using the coordinates from a dimension reduction
algorithm \(\bm{y}\), angles, distances, and volumes in this Euclidean
`output space' are not the same as on the manifold since dimension
reduction algorithms introduce distortions. To alleviate this issue,
@Perrault-Joncas2013-pq propose a method to augment
\(\bm{y}_1,\bm{y}_2\dots,\bm{y}_n\) with \(d\times d\) positive definite
matrices, \(\bm{H}_1,\bm{H}_2\dots,\bm{H}_n\), at each data point. These
matrices estimate the Riemannian metric in local coordinates defined by
the dimension reduction algorithm.

While full details are provided in @Perrault-Joncas2013-pq, we briefly
describe the Learn Metric algorithm here. There are four main steps in
the algorithm. First, a weighted neighborhood graph is constructed, with
edges between \(\bm{p}_i\) and \(\bm{p}_j\) when \(\bm{p}_i\) is a
fixed-radius nearest neighbor of \(\bm{p}_j\) or vice versa, and edge
weights depending on the distance between \(\bm{p}_i\) and \(\bm{p}_j\)
on the manifold. The fixed-radius method considers all points within a
specified radius of \(\sqrt{\epsilon}\), allowing the graph to adapt to
varying densities in the data and to capture local structure more
effectively. Second, the discrete Laplacian on this graph
\(\hat{\mathcal{L}}_{\varepsilon,n}\) is estimated {[}@Zhou2011-za{]},
where \(\sqrt{\varepsilon}\) is the radius parameter in the previous
step and a constant value \(c=0.25\) is used for the use of heat kernel
in the weighted neighborhood graph. Third, a dimension reduction method
is applied to obtain the output embedding \(\bm{y}_1,\dots,\bm{y}_n\).
Fourth, the Riemannian metric at each point is estimated by exploiting
the connection between the Riemannian metric and the Laplace Beltrami
operator (to which the graph Laplacian at Step 2 is a discrete
estimator). Full details on these four steps are provided in
\autoref{alg:learnmetric}. This algorithm is implemented in a Python
library \emph{megaman} {[}@McQueen2016-xz{]} although our own results
are based on a re-implementation of the algorithm in \emph{R}. For
computational efficiency, approximate nearest neighbor searching methods
are implemented to construct the neighborhood graph {[}@Cheng2021-dh;
@Perrault-Joncas2013-pq{]}, which are then used in both steps 1 and 3.
Two parameters, \(c=0.25\) and \(\sqrt{\varepsilon} = 0.4\), are set as
suggested in the \emph{megaman} library.

As pointed out by @Perrault-Joncas2013-pq, dimension reduction can be
carried out such that the dimension of the output vectors is larger than
the intrinsic manifold dimension \(d\). In this case, the ranks of the
matrices \(\bm{H}_i\) are equal to \(d\). Using a larger embedding
dimension is justified since it is in general not possible to embed a
manifold of dimension \(d\) globally into \(d\)-dimensional Euclidean
space. In our simulated examples, we abstract from this issue by
constructing examples that can be globally embedded into
\(d\)-dimesional Euclidean space. In practice, to determine the
dimension of the manifold, the \emph{two-nearest neighbor estimator
(TWO-NN estimator)} {[}@Facco2017-rl; @Denti2021-jl{]} can be used (See
Chapter 3 of @Lee2007-wq for more about intrinsic dimension estimation).
The \emph{R} library \emph{intRinsic} {[}@Denti2021-qc{]} implements
this algorithm and is used in all examples involving real data where the
intrinsic dimension is unknown.

\begin{algorithm}[!htb]
  \caption{Learn metric algorithm in \cite{Perrault-Joncas2013-pq} }
  \label{alg:learnmetric}
  \DontPrintSemicolon
  \SetAlgoLined
  \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}\SetKwInOut{Parameter}{parameter}\SetKwInOut{OptParameter}{optimization parameter}
  \Input{ high-dimensional data $\bm{x}_i \in \mathbb{R}^s$ for all $i=1,\ldots,N$ }
  \Output{ low-dimensional data $\bm{y}_i \in \mathbb{R}^d$ and its Riemannian metric $\bm{H}_i$ for all $i=1,\ldots,N$ }
  \Parameter{ embedding dimension $d$, bandwidth parameter $\sqrt{\varepsilon}$, manifold learning algorithm EMBED }
  \OptParameter{ manifold learning parameters in EMBED }
  \BlankLine
  \begin{algorithmic}[1]

  \STATE Construct a weighted neighborhood graph $\mathcal{G}_{w,\varepsilon}$ with weight matrix $\bm{W}$ where $w_{i,j}=\exp(-\frac{1}{\varepsilon}\|\bm{x}_i-\bm{x}_j\|^2)$ for data points $\bm{x}_i,\bm{x}_j \in \mathbb{R}^s$;

  \STATE Calculate the $N\times N$ geometric graph Laplacian $\widetilde{\mathcal{L}}_{\varepsilon,N}$ by
  $$
  \widetilde{\mathcal{L}}_{\varepsilon,N} = 1/(c\varepsilon)(\widetilde{D}^{-1} \widetilde{W} - I_N),
  $$
  where $c=0.25$, $\widetilde{D}=diag{\widetilde{W}\bm{1}}$, $\widetilde{W} = D^{-1}WD^{-1}$, and $D = diag{W\bm{1}}$;

  \STATE Embed all data point $\bm{X}\in \mathbb{R}^s$ to embedding coordinates $\bm{Y}=(\bm{y}^1,\dots,\bm{y}^d)^\prime$ by any existing manifold learning algorithm EMBED;

  \STATE Obtain the $d\times d\times N$ tensor $\bm{\tilde{H}}$ by applying the graph Laplacian $\widetilde{\mathcal{L}}_{\sqrt{\varepsilon},N}$ to the embedding coordinates matrix $\bm{Y}$ with elements in $\bm{\tilde{H}}$ found by
  $$
    \bm{\tilde{H}}^{lk \cdot} = \frac{1}{2} \left[\tilde{\mathcal{L}}_{\varepsilon, N}\left(\bm{y}^l \odot \bm{y}^k\right) - \bm{y}^l \odot\left(\tilde{\mathcal{L}}_{\varepsilon, n} \bm{y}^k\right) - \bm{y}^k \odot\left(\tilde{\mathcal{L}}_{\varepsilon, n} \bm{y}^l\right)\right],
  $$
  where $l,k=1,\dots,d$, $\bm{\tilde{H}}^{lki}$ is the element in row $l$ and column $k$ of the inverse Riemannian corresponding to observation $i$, and the $\odot$ calculation is the element-wise product between two vectors;

  \STATE For each embedding data point $\bm{y}_i, i=1,\dots,N$, calculate its $d\times d$ Riemannian metric $\bm{H}_i$ as the rank $d$ pseudo inverse of $\tilde{\bm{H}^{\cdot\cdot i}}$ below
  $$
    \bm{H}_i = \bm{U} \bm{\Lambda}_d^{-1} \bm{U}^\prime,
  $$
  where $[\bm{U}, \bm{\Lambda}]$ is the eigendecomposition of the $d\times d$ matrix $\tilde{\bm{H}^{\cdot\cdot i}}$, $\bm{U}$ is the matrix of column eigenvectors ordered by the eigenvalues in descending order, and $\bm{\Lambda}_d$ is a diagonal matrix containing the top $d$ largest eigenvalues.

  \end{algorithmic}
\end{algorithm}

\hypertarget{DCKDE}{%
\subsection{Distortion corrected KDE}\label{DCKDE}}

With all fundamentals introduced, we can now give our novel Distortion
Corrected KDE (DC-KDE) as \begin{equation}
\label{eq:denestimator}
\hat{f}(\bm{y}_j) = \frac{1}{N} \sum_{i=1}^{N} \frac{1}{r^d} \bigg(\frac{|\det \bm{H}_j|}{|\det \bm{H}_i|} \bigg)^{1/2} K\bigg( \frac{\| \bm{H}^{-1/2}_i (\bm{y}_j - \bm{y}_i)\|}{r} \bigg).
\end{equation}

The estimator has a similar structure to Equation @ref(eq:denriem) with
some key differences. To understand these differences, it is first
critical to appreciate that the coordinates
\(\bm{H}^{-1/2}_i (\bm{y}_j - \bm{y}_i)\) give an embedding that is
approximately isometric in a small neighborhood around the \(i^{th}\)
observed point (this insight is discussed at length in Section 6.2 of
@Perrault-Joncas2013-pq). This is crucial for two reasons. First, this
implies that the term \(\| \bm{H}^{-1/2}_i (\bm{y}_j - \bm{y}_i)\|\)
approximates the geodesic distance between \(\bm{y}_i\) and
\(\bm{y}_j\). Second, the estimator in Equation @ref(eq:denriem) is
valid only when the coordinate mapping is the logarithmic map around
\(\bm{y}_i\), and it is this mapping that is approximated by
\(\bm{H}^{-1/2}_i (\bm{y}_j - \bm{y}_i)\). For this reason, there is a
ratio of two determinants to ensure the density integrates to one, the
first is a consequence of the mapping from the manifold to the
coordinate system (from a dimension reduction algorithm), while the
second is the transformation \(\bm{H}^{-1/2}_i (\bm{y}_j - \bm{y}_i)\)
which ensures that the embedding approximates the logarithmic map. Also
worth noting is the resemblance between the estimator and multivariate
variable bandwidth estimation {[}@Breiman1977-qc; @Jones1990-oe;
@Terrell1992-ut{]}.

One limitation of the kernel density estimator is that the density can
be estimated only at points where data have been observed since the
estimator requires the Riemannian \(\bm{H}_j\). To estimate the density
at points that do not correspond to observed data, any smoothed average
of nearest neighbors can be used instead. We note that the particular
downstream task that we are interested in is anomaly detection for which
only the density estimates at observed sample points are required since
anomalies are identified as the points with the lowest density. The
entire workflow is summarized in \autoref{fig:vkde}. The last two steps
in \autoref{fig:vkde} are our main contributions, generating
distortion-corrected KDE with adaptive Riemannian metric \(\bm{H}_i\) at
each point and computing the highest density region plots based on the
density estimates for anomaly detection. Compared to the anomaly
detection with a general kernel density estimator in @Cheng2021-dh, the
changes are also highlighted in blue. With this anomaly detection
process, outliers based on the lowest densities could be detected more
accurately regardless of the distortion in manifold learning.

(ref:vkdecaption) The proposed schematic for anomaly detection with
distortion-corrected kernel density estimates.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{figures/DC-KDE}

}

\caption{(ref:vkdecaption)}\label{fig:vkde}
\end{figure}

\hypertarget{simulation}{%
\section{Simulations}\label{simulation}}

In this section, we examine two scenarios for both low and high
dimensions to test our proposed distortion-corrected KDE. For
visualization purposes, \autoref{twodgaussian} presents an example of a
two-dimensional manifold embedded in 3-dimensional ambient Euclidean
space. As a high-dimensional example, the second simulation in
\autoref{fivedgaussian} is based on a 4-dimensional manifold embedded in
a 100-dimensional ambient space. To estimate the density, we use the
dimension reduction algorithms ISOMAP, LLE, Laplacian Eigenmaps, t-SNE,
and UMAP. In general, we aim to highlight two advantages of our proposed
distortion-corrected KDE compared to KDE applied directly to the output
coordinates. First, the density estimates are closer to the ground truth
when distortion correction is used, and as a consequence, distortion
correction is more adept at detecting anomalies. Second, we show how
density estimation and anomaly detection are more robust to a different
choice of dimension reduction method when distortion correction is used.

\hypertarget{twodgaussian}{%
\subsection{Twin peaks example}\label{twodgaussian}}

The simulation setup for the twin peaks example is to first generate
vectors \(\bm{v}_1,\dots,\bm{v}_N\) for \(N=2000\) from a 2-dimensional
Gaussian mixture model. The mixture has four components with different
means
\(\bm{\mu_1}=(0.25, 0.25)^\prime, \bm{\mu_2}=(0.25, 0.75)^\prime, \bm{\mu_3}=(0.75, 0.25)^\prime, \bm{\mu_4}=(0.75, 0.75)^\prime\)
and the same variance-covariance matrix
\(\bm{\Sigma}_i=diag(0.016, 0.016), i=1,2,3,4\). The mixture proportions
are equally set as \(\pi_i=0.25, i=1,2,3,4\). The two-dimensional data
in \autoref{fig:metadensity} is mapped to a `twin peaks' surface via the
following \begin{equation}
\label{eq:twinpeak}
\begin{array}{lcl}
x_1 = v_1, \\
x_2 = v_2, \\
x_3 = \sin(\pi v_1) \tan (3 v_2).
\end{array}
\end{equation}

The three-dimensional twin peaks mapping is shown in
\autoref{fig:twinpeaks3d}. The colors in both \autoref{fig:metadensity}
and \autoref{fig:twinpeaks3d} indicate the true density of the data via
the twin peaks mapping, with lower density points in darker colors
scattered in the outer as well as center areas.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{figures/truedensity_twinpeaks_dc_labv}

}

\caption{Underlying data for the Gaussian mixture model of four kernels with means $(0.25, 0.25), (0.25, 0.75), (0.75, 0.25), (0.75, 0.75)$ and the same variance-covariance matrix $diag(0.016, 0.016)$. The colors indicate the true density of the data when they are mapped via the twin peaks function. Lower density points in darker colors are scattered both in the outer and center areas.}\label{fig:metadensity}
\end{figure}

(ref:twinpeaks3dcaption) Scatterplot of the 3-d twin peaks data with the
same colors indicating the true density as in \autoref{fig:metadensity}.

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{figures/scatterplot3d_twinpeaks_dc}

}

\caption{(ref:twinpeaks3dcaption)}\label{fig:twinpeaks3d}
\end{figure}

It is important to note that the \emph{true density} on the manifold is
not simply a Gaussian mixture, since the mapping in Equation
@ref(eq:twinpeak) distorts the distribution. To recover the true
distribution requires the correct Jacobian term for the pushforward from
\(\bm{v}\) to the volume form of twin peaks manifold. By treating the
\(\bm{v}\) as an `output' embedding from input points \(\bm{x}\) that
lie on the true manifold and applying the Learn Metric algorithm, we can
obtain \(\bm{\Gamma}_i\) for \(i=1,\dots,n\) where \(\bm{\Gamma}_i\) is
the Riemannian metric of the coordinate system given by \(\bm{v}\). This
notation is distinct from \(\bm{H}_i\) which is the output of the Learn
Metric algorithm for a coordinate system obtained via a dimension
reduction algorithm. The true density on the manifold can be obtained as
\(f(\bm{p}_i)=f(\bm{v}_i)|\bm{\Gamma}_i|^{1/2}\), where \(f(\bm{v}_i)\)
is the density of a four-component mixture of normals. Knowledge of
\(\bm{v}\) and \(\bm{\Gamma_i}\) will not be used when estimating the
density but only to establish a `ground truth' for densities on the
manifold. \autoref{fig:metadensity} shows the simulated \(\bm{v}\) with
color indicating the true density of data on the manifold. Anomalies are
defined as points with the lowest densities shown in darker colors and
with `typical' points having the highest density shown in yellow. The
anomalies are found around the edges of the plot, but there is also a
low-density region between the means of the four mixture components. The
objective is to determine whether we can correctly identify these
features without any knowledge of the true density or the \(\bm{v}\).

\begin{figure}[!p]

{\centering \includegraphics[width=0.85\linewidth]{figures/Twin Peak2000_5levels_outliers_comparison_5ml_r0_5}

}

\caption{Highest density region plots of five manifold learning embeddings of the twin peaks data in each row. The top 20 outliers, highlighted in black and indexed in blue text, are found by the true manifold density (left panel), DC-KDE (middle panel), and KDE (right panel). DC-KDE finds more true outliers than KDE in all five rows.}\label{fig:tpoutliers}
\end{figure}

\autoref{fig:tpoutliers} summarizes the results. Each row of panels
corresponds to a different dimension reduction technique, while the
left, center and right columns correspond to density estimates for the
ground truth density, distortion-corrected KDE, and KDE respectively. We
set the bandwidth parameter in the DC-KDE @ref(eq:denestimator) as
\(r=0.5\). Colors show the different estimated densities at each point
with anomalies shown in black with blue indexing, and higher density
points shown in yellow. For many methods, the salient features of the
ground truth distribution are clear regardless of whether distortion
correction is applied, for example for ISOMAP, all three plots, identify
a similar set of outliers and four high-density regions. On the other
hand for LLE, the left panel shows that dimension reduction pulls
outliers on the manifold towards the center. The distortion-corrected
KDE can account for this, while KDE without distortion correction on the
other hand does not correctly identify the anomalies. For t-SNE, the
ground truth and distortion-corrected KDE identify four regions of high
density, while a KDE estimate without distortion correction seems to
identify a larger number of modes. This concurs with the common
observation that t-SNE tends to output clusters even where such clusters
may not be present in the underlying data {[}@cai2022-tsne{]}.

\begin{table}

\caption{\label{tab:tpcors}Correlation between true density ranking and estimated density ranking for different manifold learning embeddings of the twin peak data. Distortion-corrected KDE outperforms for all dimension reduction algorithms and gives the higher rank correlation to the output of t-SNE and UMAP.}
\centering
\begin{tabular}[t]{l>{}l>{}l>{}l>{}l>{}l}
\toprule
  & ISOMAP & LLE & Laplacian.Eigenmaps & t.SNE & UMAP\\
\midrule
DC-KDE & \textbf{0.823} & \textbf{0.673} & \textbf{0.672} & \textbf{0.806} & \textbf{0.794}\\
KDE & 0.798 & 0.500 & 0.606 & 0.451 & 0.469\\
\bottomrule
\end{tabular}
\end{table}

We can gain further insight by comparing the correlation between ranks
of true densities and estimated densities from KDE with and without
density correction by \autoref{tab:tpcors}. Distortion correction
improves the rank correlation for all dimension reduction algorithms. In
particular, while applying KDE to the output of t-SNE and UMAP leads to
a moderate correlation below 0.5, applying distortion correction
improves these rank correlations to values close to 0.8.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/Twin Peak2000_densityrank_comparison_isomap_radius8_r0_5_logrank_rec_colprob_smallblocks3_crossfalse}

}

\caption{Scatterplot of log scale ranks of true density and estimated density ranks for DC-KDE (in the left panel) and KDE (in the right panel) based on ISOMAP embedding. The colors indicate different levels of highest density regions and the shapes indicate whether the density estimators correctly classify the true anomalies. The shading contains all anomalies that are both truly and correctly identified. KDE without distortion correction gives more misclassified anomalies.}\label{fig:tpisomapden}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/Twin Peak2000_densityrank_comparison_tsne_radius8_r0_5_logrank_rec_colprob_smallblocks3_crossfalse}

}

\caption{Scatterplot of log scale ranks of true density and estimated density ranks for DC-KDE (in the left panel) and KDE (in the right panel) based on t-SNE embedding. KDE without distortion correction gives many more misclassified anomalies.}\label{fig:tptsneden}
\end{figure}

In \autoref{fig:tpisomapden}, we plot ranks of the estimated density
against the true density for the ISOMAP embedding with the left panels
showing results for distortion correction and the right panel showing
results without distortion correction. Data are presented on a log scale
to highlight anomalies. The bottom left shaded region contains all
points that are truly anomalies and are identified as such (true
positives), where an anomaly is defined as a point not falling within a
\(99\%\) highest density region. The middle shaded region contains
anomalies that are true positives in the sense of not lying in a
\(90\%\) HDR. Points lying outside squares (shown as triangles) are
incorrectly classified. For example, the three red triangles in the
middle left of the left panel are truly anomalies since they lie outside
the \(99\%\) HDR, but are not classified as such (although they are true
positives if a 90\% HDR is used). Overall, the right panel contains many
more misclassified anomalies, which shows that failing to apply
distortion correction can have a severe impact on anomaly detection.
\autoref{fig:tptsneden} shows the same plot but for t-SNE. The quality
of t-SNE is worse than ISOMAP in this example therefore many more
anomalies are misclassified. However, the difference between KDE with
and without distortion correction is stark. These results highlight the
importance of applying distortion correction especially when the quality
of dimension reduction may not be high.

\begin{figure}[!p]

{\centering \includegraphics[width=0.8\linewidth]{figures/Twin Peak2000_density_compare_isomapvs4ml_radius8_r0_5_rank}

}

\caption{Comparison of ranks of the estimated densities based on ISOMAP  and four other dimension reduction algorithms in each row. Distortion-corrected KDE (on the left panel) and KDE (on the right panel) are compared and DC-KDE shows the robustness to the use of different dimension reduction methods.}\label{fig:tpisomapvs4ml}
\end{figure}

\FloatBarrier

Finally, \autoref{fig:tpisomapvs4ml} demonstrates the robustness of
distortion correction methods to the use of a dimension reduction
algorithm. Each row of panels compares ranks of the estimated densities
based on a dimension reduction algorithm to the estimated density based
on ISOMAP. The left column shows results when distortion correction is
applied, and the right column when it is not applied. It can be seen
that the rank correlation between estimates based on different dimension
reduction algorithms is much higher when distortion correction is
applied. This is critical since conclusions will be more robust to the
choice of dimension reduction algorithm.

\hypertarget{fivedgaussian}{%
\subsection{Semi-hypersphere example embedded in 100-D
space}\label{fivedgaussian}}

As a high-dimensional experiment, we generate the underlying data from a
5-dimensional semi-hypersphere, embedded within 100-dimensional ambient
space. To start with, we simulate vectors
\((\bm{v}_1, \dots, \bm{v}_N)^\prime\) for \(N=10,000\) points from a
4-dimensional Gaussian mixture model with two mixture components,
\(\mathcal{N}(\bm{\mu}_1, \bm{\Sigma}_1)\) and
\(\mathcal{N}(\bm{\mu}_2, \bm{\Sigma}_2)\), with the same means
\(\bm{\mu}_1 = \bm{\mu}_2 =(0, 0, 0, 0)^\prime\) and different
variance-covariance matrices \(\bm{\Sigma}_1 = diag(1,1,1,1)\) and
\(\bm{\Sigma}_2 = diag(2,2,2,2)\). The mixture proportions are set as
\(\pi_1=0.99\) and \(\pi_2=0.01\). With this design, the observations
from the second component tend to be outlying anomalies. The data are
mapped to a hemisphere via the equation
\(v_1^2 + v_2^2 + v_3^2 + v_4^2 + v_5^2 = r^2\) where \(v_5>0\) and
\(r\) is set as \(8\). \autoref{fig:fivedmeta} shows a scatterplot which
is a single frame from a 5-D tour path\footnote{See the animation of the
  5-D grand tour at
  \url{https://github.com/ffancheng/kderm/blob/master/monash/figures/tourr_5d_animation.gif}.}
animation using the R package \emph{tourr} {[}@Wickham2011-ir{]}. The
round and triangular point shapes indicate the two mixture components
\(\mathcal{N}(\bm{\mu}_1, \bm{\Sigma}_1)\) and
\(\mathcal{N}(\bm{\mu}_2, \bm{\sigma}_2)\), and the colors indicate the
distance between each point and the center of the distribution. It can
be seen that the most distant points are in darker colors and triangular
shapes, meaning that the most anomalous observations are generated from
the second mixture component.

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{figures/tourr_5d_semisphere}

}

\caption{Scatterplot display of the animation of a 5-D tour path with shapes indexing two Gaussian mixture components and the colors showing the distance to the kernel cores. Distant points in darker colors could be seen as anomalies.}\label{fig:fivedmeta}
\end{figure}

To embed the 5-D hyper-semisphere into 100-D space, we append 95 zero
columns to \(\bm{v}_i\) so that
\(\bm{v}_i=(v_1, \dots, v_5, 0, \dots, 0), i=1,\dots,N\). Next, we
rotate the \(100\)-dimensional vectors
\((\bm{v}_1, \dots, \bm{v}_N)^\prime\) by multiplying by a randomly
generated rotation matrix. To generate the rotation matrix we first
simulate elements from a uniform \((0,1)\) distribution, stack them into
a \(100\times 100\) matrix \(\bm{A}\) and then take the R matrix from
the QR decomposition of \(\bm{A}\). Rotating the vectors results in
input vectors that are no longer sparse. Nonetheless, the intrinsic
dimension of this is still \(d=4\).

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/sim4d10000_density_comparison_4ml_radius10_k200_rankdensity_circleoutlier_with1rec}

}

\caption{Rank comparison between the true density and estimated density from both DC-KDE and KDE. Four manifold learning methods are used row-wise. The point shapes indicate whether they are the true outliers, and the grey shading highlights the top 1\% rank region. The colors show the distance to the center of the semisphere, with darker points being distant from the center.}\label{fig:fivedisomapden}
\end{figure}

Following similar steps to \autoref{twodgaussian}, we estimate the
densities of the 4-D manifold and compare them with the ground truth
density. The bandwidth parameter for the DC-KDE estimator is set as
\(r=1\) for this example. In \autoref{tab:fivedcors}, the rank
correlations between true densities and estimated densities from DC-KDE
and KDE are presented. For ISOMAP and LLE, KDE without distortion
correction has a slightly higher rank correlation with the true density
than for KDE with distortion correction, but these differences are
negligible and both density estimators have a very high correlation of
more than 0.96. However, for Laplacian Eigenmaps and UMAP, more
distortion is induced through dimension reduction, the rank correlation
between estimated and true densities is close to 0 when distortion
correction is not applied. In the same settings, the corresponding rank
correlations are relatively high at 0.87 and 0.78 for Laplacian
Eigenmaps and UMAP respectively, when distortion correction is applied.

\autoref{fig:fivedisomapden} is constructed in a similar fashion to
\autoref{fig:tpisomapden} to show the effect of distortion correction on
the detection of anomalies. As in \autoref{fig:fivedmeta}, the point
shapes indicate which of the two mixture Gaussian mixture components an
observation was generated from and the colors indicate the distance from
the center of the distribution with outliers shown in darker colors. The
ranks of the estimated densities are shown against ranks of true
densities (on the log scale) with panels on the left showing
distortion-corrected KDE and panels on the right showing KDE without
distortion correction. The panels from top to bottom show four dimension
reduction algorithms: ISOMAP, LLE, Laplacian Eigenmaps, and UMAP. Note
that we exclude t-SNE algorithm in this section because it is designed
mainly for low-dimensional visualization purposes and is only applicable
to an embedding dimension less than or equal to three.

Comparing the left and right panels for each row, we notice that KDE
with distortion correction has fewer misclassified observations, and
therefore outperforms KDE. For UMAP, almost all ground truth anomalies
(points outside a \(99\%\) HDR) are not correctly detected using KDE,
while almost all anomalies are correctly detected after correcting for
distortion. For DC-KDE in the left panel there are several triangles in
a horizontal line, this arises, due to many points having an estimated
density of close to zero (values less than \(10^{-7}\)) which leads to
tied ranks. These low-density points are all detected as anomalies for
all dimension reduction algorithms, which again shows the robustness of
DC-KDE.

\begin{table}

\caption{\label{tab:fivedcors}Correlation between true density and estimated density for four manifold learning embeddings of the 100-D data. DC-KDE has a slightly lower rank correlation for ISOMAP and LLE above 0.96, but significantly outperfoms KDE for Laplacian Eigenmaps and UMAP when more distortion is induced in dimension reduction.}
\centering
\begin{tabular}[t]{l>{}l>{}l>{}l>{}l}
\toprule
  & ISOMAP & LLE & Laplacian.Eigenmaps & UMAP\\
\midrule
DC-KDE & 0.968 & 0.970 & \textbf{0.8674} & \textbf{0.782}\\
KDE & \textbf{0.976} & \textbf{0.971} & 0.0328 & -0.181\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:fourdhdrtable}Proportion comparison of outliers in correct highest density regions in density estimation of four manifold learning embeddings for the 100-D data.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{l|>{}c>{}c|>{}c>{}c|>{}c>{}c|>{}cc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{ISOMAP} & \multicolumn{2}{c}{LLE} & \multicolumn{2}{c}{Laplacian Eigenmaps} & \multicolumn{2}{c}{UMAP} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
  & DC-KDE & KDE & DC-KDE & KDE & DC-KDE & KDE & DC-KDE & KDE\\
\midrule
>99\% HDR & \cellcolor[HTML]{E6E6E6}{\textbf{0.830}} & 0.490 & \cellcolor[HTML]{E6E6E6}{\textbf{0.830}} & 0.240 & \cellcolor[HTML]{E6E6E6}{\textbf{0.840}} & 0.840 & \cellcolor[HTML]{E6E6E6}{\textbf{0.770}} & 0.000\\
99\% HDR & \cellcolor[HTML]{E6E6E6}{\textbf{0.818}} & 0.685 & \cellcolor[HTML]{E6E6E6}{\textbf{0.805}} & 0.478 & \cellcolor[HTML]{E6E6E6}{\textbf{0.815}} & 0.595 & \cellcolor[HTML]{E6E6E6}{\textbf{0.632}} & 0.032\\
\bottomrule
\end{tabular}}
\end{table}

\autoref{tab:fourdhdrtable} further illustrates the proportions of
correctly classified anomalies in the \(>99\%\) and \(99\%\) highest
density regions. Compared with KDE, the proportions are always higher
when distortion correction methods are used. The proportions with
distortion correction are very close for ISOMAP, LLE, and Laplacian
Eigenmaps with a value of around \(83\%\) while slightly lower for UMAP
at around \(77\%\). This could be due to the severe distortion usually
induced by the UMAP algorithm.

\hypertarget{dckdeapplication}{%
\section{Application}\label{dckdeapplication}}

In this application, we use the smart meter data from the \emph{CER
Smart Metering Project - Electricity Customer Behaviour Trial,
2009-2010} in Ireland {[}@cer2012-data{]} between 14 July 2009 and 31
December 2010. The CER dataset\footnote{accessed via the Irish Social
  Science Data Archive - www.ucd.ie/issda.} records the half-hourly
electricity consumption of individual residential and commercial
properties, but not including energy for cooling or heating systems. We
selected the \(3,639\) residential data with no missing values during
the data collection period for a total of \(535\) days.

For the electricity consumption data of residential individuals, it is
worthwhile to explore the distribution of electricity demand rather than
the raw consumption data, so as to study the usage patterns of different
households or different periods of the week {[}@Hyndman2018-ia{]}. This
can be considered as a case of dimension reduction on a statistical
manifold, that is a manifold with elements that are probability
distributions. @Cheng2021-dh propose estimators of the Total Variation
Metric and Hellinger distance between distributions that can be used in
a computationally practical manner for dimension reduction on
statistical manifolds. Again, we use ISOMAP, LLE, Laplacian Eigenmaps,
t-SNE, and UMAP for dimension reduction to obtain a \(2\)-dimensional
embedding for kernel density estimation and anomaly detection. To this
end, we compare the density estimates from KDE with and without
distortion correction and show how robust the anomalies are to different
dimension reduction algorithms, only when distortion correction is used.
We use the highest density regions plot to visualize the density
estimates. However, for this real data set with an unknown structure,
the ground truth densities are unknown and it is not possible to tell
which anomalies are the true ones.

Although full details for data processing and dimension reduction are
provided in @Cheng2021-dh, we briefly describe the process here. For
each household, a discrete approximation of the distribution of
electricity demand at each one of the \(336\) half-hourly periods of the
week is found. For any pair of households the total variation metric can
be found between the distributions corresponding to any half hour of the
week, and summing over these gives a distance measure between the pair
of households, subsequently used for dimension reduction algorithms and
the Learn Metric algorithm. The number of nearest neighbors in manifold
learning \(K\) is set as \(100\) and the bandwidth parameter \(r=180\)
to match the distances between the distributions of households.

\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{figures/Electricity_2d_5levels_outliers_comparison_5ml_r180}

}

\caption{Highest density region plots of the 2-D embedding for $3,639$ households, with each point representing the distribution of one household and colors indicating the density estimated from DC-KDE (left panel) and KDE (right panel). The black points with blue text indexing are the top 20 anomalous households found by each dimension reduction method in each row.}\label{fig:electricityembedding}
\end{figure}
\begin{table}

\caption{\label{tab:electricityoutliers}Top 20 anomalous household IDs using density estimates of different 2-dimensional embedding. The left five columns are anomalies by DC-KDE, while the right five columns are anomalies using KDE. The blue-colored IDs are IDs that are detected over three times, with the darker one indicating five times and the lighter one indicating four times.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabu} to \linewidth {>{}c>{}c>{}c>{}c>{}c>{}c|>{}c>{}c>{}c>{}c>{}c}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{DC-KDE} & \multicolumn{5}{c}{KDE} \\
\cmidrule(l{3pt}r{3pt}){2-6} \cmidrule(l{3pt}r{3pt}){7-11}
  & ISOMAP & LLE & Laplacian Eigenmaps & t-SNE & UMAP & ISOMAP & LLE & Laplacian Eigenmaps & t-SNE & UMAP\\
\midrule
1 & \textcolor[HTML]{000000}{6317} & \textcolor[HTML]{2171B5}{\textbf{7049}} & \textcolor[HTML]{2171B5}{\textbf{3161}} & \textcolor[HTML]{000000}{4826} & \textcolor[HTML]{2171B5}{\textbf{4546}} & \textcolor[HTML]{000000}{6317} & \textcolor[HTML]{000000}{1881} & \textcolor[HTML]{000000}{6317} & \textcolor[HTML]{000000}{4190} & \textcolor[HTML]{000000}{6533}\\
2 & \textcolor[HTML]{000000}{1353} & \textcolor[HTML]{2171B5}{\textbf{4546}} & \textcolor[HTML]{000000}{5269} & \textcolor[HTML]{000000}{1069} & \textcolor[HTML]{2171B5}{\textbf{7049}} & \textcolor[HTML]{000000}{4884} & \textcolor[HTML]{6BAED6}{\textbf{6199}} & \textcolor[HTML]{000000}{6643} & \textcolor[HTML]{000000}{6817} & \textcolor[HTML]{000000}{5899}\\
3 & \textcolor[HTML]{000000}{4884} & \textcolor[HTML]{000000}{1881} & \textcolor[HTML]{000000}{4875} & \textcolor[HTML]{2171B5}{\textbf{4546}} & \textcolor[HTML]{000000}{1069} & \textcolor[HTML]{000000}{5092} & \textcolor[HTML]{6BAED6}{\textbf{5356}} & \textcolor[HTML]{000000}{6764} & \textcolor[HTML]{000000}{6492} & \textcolor[HTML]{000000}{6345}\\
4 & \textcolor[HTML]{2171B5}{\textbf{4546}} & \textcolor[HTML]{000000}{6952} & \textcolor[HTML]{000000}{7300} & \textcolor[HTML]{2171B5}{\textbf{7049}} & \textcolor[HTML]{000000}{4862} & \textcolor[HTML]{000000}{1353} & \textcolor[HTML]{000000}{4233} & \textcolor[HTML]{000000}{1058} & \textcolor[HTML]{000000}{2309} & \textcolor[HTML]{000000}{1884}\\
5 & \textcolor[HTML]{2171B5}{\textbf{7049}} & \textcolor[HTML]{000000}{6838} & \textcolor[HTML]{000000}{4880} & \textcolor[HTML]{6BAED6}{\textbf{1533}} & \textcolor[HTML]{000000}{3639} & \textcolor[HTML]{000000}{5580} & \textcolor[HTML]{2171B5}{\textbf{7049}} & \textcolor[HTML]{000000}{5105} & \textcolor[HTML]{000000}{2546} & \textcolor[HTML]{000000}{3278}\\
\addlinespace
6 & \textcolor[HTML]{000000}{1366} & \textcolor[HTML]{000000}{3874} & \textcolor[HTML]{000000}{5599} & \textcolor[HTML]{000000}{4265} & \textcolor[HTML]{6BAED6}{\textbf{1964}} & \textcolor[HTML]{000000}{5594} & \textcolor[HTML]{000000}{1821} & \textcolor[HTML]{000000}{6577} & \textcolor[HTML]{000000}{2907} & \textcolor[HTML]{000000}{5463}\\
7 & \textcolor[HTML]{2171B5}{\textbf{3161}} & \textcolor[HTML]{000000}{1069} & \textcolor[HTML]{000000}{3222} & \textcolor[HTML]{000000}{4862} & \textcolor[HTML]{000000}{2249} & \textcolor[HTML]{000000}{2576} & \textcolor[HTML]{2171B5}{\textbf{4546}} & \textcolor[HTML]{000000}{3880} & \textcolor[HTML]{000000}{4536} & \textcolor[HTML]{000000}{5504}\\
8 & \textcolor[HTML]{000000}{1074} & \textcolor[HTML]{000000}{2249} & \textcolor[HTML]{000000}{1776} & \textcolor[HTML]{000000}{3639} & \textcolor[HTML]{6BAED6}{\textbf{6199}} & \textcolor[HTML]{000000}{1366} & \textcolor[HTML]{000000}{3431} & \textcolor[HTML]{000000}{5358} & \textcolor[HTML]{000000}{6898} & \textcolor[HTML]{000000}{1782}\\
9 & \textcolor[HTML]{000000}{6681} & \textcolor[HTML]{6BAED6}{\textbf{1964}} & \textcolor[HTML]{000000}{1161} & \textcolor[HTML]{000000}{5546} & \textcolor[HTML]{000000}{3892} & \textcolor[HTML]{000000}{6681} & \textcolor[HTML]{000000}{1569} & \textcolor[HTML]{000000}{1366} & \textcolor[HTML]{000000}{4128} & \textcolor[HTML]{000000}{5965}\\
10 & \textcolor[HTML]{000000}{1701} & \textcolor[HTML]{000000}{1569} & \textcolor[HTML]{000000}{3655} & \textcolor[HTML]{000000}{4896} & \textcolor[HTML]{000000}{6209} & \textcolor[HTML]{000000}{5291} & \textcolor[HTML]{6BAED6}{\textbf{1964}} & \textcolor[HTML]{000000}{5545} & \textcolor[HTML]{000000}{7368} & \textcolor[HTML]{000000}{6081}\\
\addlinespace
11 & \textcolor[HTML]{000000}{3136} & \textcolor[HTML]{6BAED6}{\textbf{1533}} & \textcolor[HTML]{000000}{6169} & \textcolor[HTML]{000000}{4555} & \textcolor[HTML]{000000}{6952} & \textcolor[HTML]{2171B5}{\textbf{3161}} & \textcolor[HTML]{000000}{6568} & \textcolor[HTML]{000000}{2160} & \textcolor[HTML]{000000}{2232} & \textcolor[HTML]{000000}{4894}\\
12 & \textcolor[HTML]{000000}{1881} & \textcolor[HTML]{000000}{4555} & \textcolor[HTML]{000000}{4830} & \textcolor[HTML]{6BAED6}{\textbf{6199}} & \textcolor[HTML]{000000}{4896} & \textcolor[HTML]{000000}{1701} & \textcolor[HTML]{000000}{7300} & \textcolor[HTML]{000000}{6585} & \textcolor[HTML]{000000}{6273} & \textcolor[HTML]{000000}{5937}\\
13 & \textcolor[HTML]{000000}{6785} & \textcolor[HTML]{000000}{6526} & \textcolor[HTML]{000000}{2284} & \textcolor[HTML]{000000}{6952} & \textcolor[HTML]{000000}{6838} & \textcolor[HTML]{000000}{4749} & \textcolor[HTML]{000000}{7438} & \textcolor[HTML]{000000}{5824} & \textcolor[HTML]{000000}{1557} & \textcolor[HTML]{000000}{5522}\\
14 & \textcolor[HTML]{000000}{5092} & \textcolor[HTML]{000000}{3639} & \textcolor[HTML]{000000}{6526} & \textcolor[HTML]{000000}{6838} & \textcolor[HTML]{6BAED6}{\textbf{1533}} & \textcolor[HTML]{000000}{6699} & \textcolor[HTML]{000000}{4401} & \textcolor[HTML]{000000}{4932} & \textcolor[HTML]{000000}{3500} & \textcolor[HTML]{000000}{5580}\\
15 & \textcolor[HTML]{000000}{5269} & \textcolor[HTML]{6BAED6}{\textbf{6199}} & \textcolor[HTML]{000000}{1694} & \textcolor[HTML]{6BAED6}{\textbf{1964}} & \textcolor[HTML]{000000}{1569} & \textcolor[HTML]{000000}{5504} & \textcolor[HTML]{000000}{3434} & \textcolor[HTML]{000000}{6410} & \textcolor[HTML]{000000}{1932} & \textcolor[HTML]{000000}{6902}\\
\addlinespace
16 & \textcolor[HTML]{000000}{2576} & \textcolor[HTML]{000000}{5531} & \textcolor[HTML]{000000}{3431} & \textcolor[HTML]{000000}{7104} & \textcolor[HTML]{000000}{3213} & \textcolor[HTML]{000000}{3170} & \textcolor[HTML]{000000}{2249} & \textcolor[HTML]{000000}{3023} & \textcolor[HTML]{000000}{3119} & \textcolor[HTML]{000000}{7243}\\
17 & \textcolor[HTML]{000000}{1262} & \textcolor[HTML]{000000}{4875} & \textcolor[HTML]{000000}{2523} & \textcolor[HTML]{000000}{3892} & \textcolor[HTML]{000000}{4401} & \textcolor[HTML]{000000}{5899} & \textcolor[HTML]{000000}{3213} & \textcolor[HTML]{000000}{1497} & \textcolor[HTML]{000000}{3809} & \textcolor[HTML]{000000}{6699}\\
18 & \textcolor[HTML]{000000}{5966} & \textcolor[HTML]{6BAED6}{\textbf{5356}} & \textcolor[HTML]{6BAED6}{\textbf{5356}} & \textcolor[HTML]{2171B5}{\textbf{3161}} & \textcolor[HTML]{000000}{6766} & \textcolor[HTML]{000000}{2658} & \textcolor[HTML]{2171B5}{\textbf{3161}} & \textcolor[HTML]{000000}{7293} & \textcolor[HTML]{000000}{6108} & \textcolor[HTML]{000000}{5549}\\
19 & \textcolor[HTML]{000000}{5937} & \textcolor[HTML]{000000}{3892} & \textcolor[HTML]{000000}{2233} & \textcolor[HTML]{000000}{6799} & \textcolor[HTML]{000000}{4265} & \textcolor[HTML]{000000}{5189} & \textcolor[HTML]{6BAED6}{\textbf{1533}} & \textcolor[HTML]{000000}{5580} & \textcolor[HTML]{000000}{4516} & \textcolor[HTML]{000000}{5048}\\
20 & \textcolor[HTML]{6BAED6}{\textbf{5356}} & \textcolor[HTML]{000000}{4896} & \textcolor[HTML]{000000}{5125} & \textcolor[HTML]{000000}{6209} & \textcolor[HTML]{000000}{6799} & \textcolor[HTML]{000000}{2669} & \textcolor[HTML]{000000}{4555} & \textcolor[HTML]{000000}{6224} & \textcolor[HTML]{000000}{5937} & \textcolor[HTML]{000000}{1923}\\
\bottomrule
\end{tabu}}
\end{table}

\autoref{fig:electricityembedding} presents the highest density regions
of the 2-dimensional embedding for each dimension reduction method,
ISOMAP, LLE, Laplacian Eigenmaps, t-SNE, and UMAP, with the left panels
showing the densities estimated with the distortion corrected KDE and
the right panels showing those estimate with KDE. There are six density
regions colored in the plot, \(1\%\), \(5\%\), \(50\%\), \(90\%\),
\(99\%\) and \(>99\%\), and darker colors indicate lower density
regions. The black points indexed with blue household IDs are the top 20
anomalies with the lowest densities. First of all, for the same
electricity data, each dimension reduction method generates quite
different embeddings, meaning that distortions exist in these
embeddings. When comparing row-wise for the same embedding, it can be
seen that ISOMAP and LLE give similar anomalies while the outliers are
more clustered for Laplacian Eigenmaps, t-SNE, and UMAP with DC-KDE
compared with the KDE. As the anomalous IDs are not easy to distinguish
in the plots, we list them in \autoref{tab:electricityoutliers} and
highlight the IDs that are most identified as anomalies. The darker blue
points are those households that are identified five times out of the
ten cases shown in \autoref{fig:electricityembedding}, and lighter blue
points are those identified four times. Again, KDE with distortion
correction methods finds more anomalies in common regardless of the
dimension reduction methods, compared with KDE where only LLE could find
similar anomalies to other cases. Without distortion correction, the
density estimates are different not only in scale, but also show
inconsistency in the rank of density estimates.

\begin{table}

\caption{\label{tab:electricityhdrpercent}Proportion of the same outliers as UMAP in two highest density regions when estimating densities for four manifold learning embeddings.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{l|>{}c>{}c|>{}c>{}c|>{}c>{}c|>{}cc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{ISOMAP} & \multicolumn{2}{c}{LLE} & \multicolumn{2}{c}{Laplacian Eigenmaps} & \multicolumn{2}{c}{t-SNE} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
  & DC-KDE & KDE & DC-KDE & KDE & DC-KDE & KDE & DC-KDE & KDE\\
\midrule
>99\% HDR & \cellcolor[HTML]{E6E6E6}{0.189} & \textbf{0.243} & \cellcolor[HTML]{E6E6E6}{\textbf{0.189}} & 0.027 & \cellcolor[HTML]{E6E6E6}{\textbf{0.622}} & 0.000 & \cellcolor[HTML]{E6E6E6}{\textbf{0.865}} & 0.108\\
99\% HDR & \cellcolor[HTML]{E6E6E6}{\textbf{0.359}} & 0.248 & \cellcolor[HTML]{E6E6E6}{\textbf{0.503}} & 0.000 & \cellcolor[HTML]{E6E6E6}{\textbf{0.697}} & 0.007 & \cellcolor[HTML]{E6E6E6}{\textbf{0.628}} & 0.083\\
\bottomrule
\end{tabular}}
\end{table}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/Electricity_2d_rankplot_3639id_umapvs4ml_radius20_r180}

}

\caption{Comparison of ranks of the estimated densities based on UMAP and four dimension reduction algorithms in each row. Distortion-corrected KDE (on the left panel) and KDE (on the right panel) are compared and DC-KDE shows a higher level of robustness to the use of different dimension reduction methods.}\label{fig:electricityrank}
\end{figure}

Unlike for the simulations where the ground truth densities are known,
in the applications, we have no knowledge of the truly anomalous
households. To explore the robustness of anomaly detection both with and
without distortion correction, we assume, in this case, that the
anomalies given by UMAP are the `true' ones and evaluate the performance
of the other four dimension reduction methods, ISOMAP, LLE, Laplacian
Eigenmaps and t-SNE. The results are summarized in
\autoref{tab:electricityhdrpercent} which can be interpreted similarly
to \autoref{tab:fourdhdrtable}. \autoref{tab:electricityhdrpercent}
presents the proportions of `true' outliers (given by UMAP) where
anomalies fall outside the \(>99\%\) and \(99\%\) HDRs respectively that
are also identified as anomalies by different dimension reduction
methods. Both DC-KDE and KDE are shown in the table. Apart from the
lower percentage of ISOMAP in \(>99\%\) HDRs, distortion-corrected KDE
gives higher proportions of true outliers in all other methods. The
improvement brought by distortion correction is much more pronounced in
Laplacian Eigenmaps and t-SNE which would otherwise detect almost
completely different anomalies to UMAP. \autoref{fig:electricityrank}
presents the visualization of density ranking in log scale. The grey
shading area covers the density rank from 1 to 20, which are the
anomalies listed in \autoref{tab:electricityoutliers}. Again, the plots
show the robustness of density estimated with distortion correction
regardless of the dimension reduction methods. The robustness of anomaly
detection to different methods when using distortion correction is also
shown in \autoref{tab:electricityoutliers}. The household indices shown
in darker shades of blue are identified as anomalies the most with
dimension reduction algorithms. It is notable that without applying
distortion correction, Laplacian Eigenmaps, t-SNE and UMAP do not
identify any of the same anomalies as ISOMAP and LLE.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{figures/Electricity_2d_compare4id_1typical_3anomalous}

}

\caption{Electricity usage plots of all 535 days for the most typical household (top row) and three anomalies.}\label{fig:electricityraw}
\end{figure}

We now further examine households 3161, 4546, and 7049, which appear as
anomalies for at least three dimension reduction algorithms as well as
Households 2396, which is the most typical household with the highest
density estimate. Their half-hourly electricity demands across the
entire trial period are plotted in \autoref{fig:electricityraw}.
Compared with ID 2396, the other three households show different usage
patterns. ID 3161 has much less electricity usage in summers, mostly
below 1kWh, than in winters, above 3kWh. As this data set does not
include the energy used for cooling or heating systems, it is possible
that there are other large electronic appliances used in winter for this
household. The other two anomalous IDs, 4546 and 7049, however, have
nearly zero kWh electricity consumption for many days and some spikes
for other days. It is likely that these two households are not occupied
all the time during the trial and could be used as temporary
accommodation (e.g.~through a holiday rental service such as Airbnb).

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{figures/Electricity_2d_plot_hdrbox_7dow_4id_1typical_3anomalous}

}

\caption{Quantile region plots of electricity demand against the time of the week for one typical household 1472 and three anomalies, 3161, 4546, and 7049. The day-of-the-week patterns are much more different for Household 4546 and 7049 compared with 3161.}\label{fig:electricityhdrsvkde}
\end{figure}

Further insights could be gained by comparing the quantile region plots
of electricity demand against the time of the week for the same typical
or anomalous households in \autoref{fig:electricityraw}. The
distribution of the typical households in the top panel has shown a
repeated period-of-the-week usage pattern, with higher usage during
mealtime on all seven days of the week and slightly higher usage for
weekends. This repeated pattern in a week window is also shown on some
days of the week for household ID 3161 with small spikes at noon and
midnight but not on Fridays and Saturdays. As for the bottom two
households, the middle row in \autoref{fig:electricityhdrsvkde} shows a
distribution with zero median and a \(75\%\) quantile of 1kWh from
around 9 pm to 6 am, while the distributions for ID 7049 do not have an
obvious pattern across different periods of the week. These findings
show that distortion correction methods work well in estimating the
densities of the electricity usage and detecting anomalous households in
the smart meter data despite the various embeddings from different
dimension reduction methods.

\hypertarget{dckdeconclusion}{%
\section{Conclusions}\label{dckdeconclusion}}

In this paper, we propose a novel distortion correction method to
estimate the density of an embedding from manifold learning algorithms
and further identify outliers based on the densities. Compared with KDE,
our distortion-corrected KDE makes use of geometric information for each
data point to correct the distortion induced by the embedding. The
Riemannian metric is estimated with the Learn Metric algorithm to
approximate the geodesic distance and volume density function locally at
each point. We compare our proposed method with KDE by two simulation
settings, a 2-D manifold embedded as a 3-D twin peaks shape and a 4-D
manifold mapped in a 100-D ambient space, and show that DC-KDE could
generate more accurate kernel density estimation is more robust to the
choice of dimension reduction algorithm.

As an empirical example, we explore the distributions of \(3,639\)
households and \(336\) time periods of the week in the Irish smart meter
data. Five manifold learning algorithms, including ISOMAP, LLE,
Laplacian Eigenmaps, t-SNE, and UMAP, are applied to get the 2-D
embedding, followed by density estimation with KDE and DC-KDE. Without
the ground truth density, we assume that UMAP gives the true densities
and compare both density estimation methods by looking at the
distributions of the most typical households with the highest densities
and three anomalous households that are identified most times. The
distortion correction methods could identify the typical households with
repeated time-of-the-week usage patterns, while the distributions of the
outliers are anomalous in different ways.

There are several open questions to be explored. The first involves the
selection of bandwidth parameters and kernel types for kernel density
estimation, which has been widely explored in many KDE-related
literature. For densities on unknown manifolds, @Berenfeld2021-as also
provide some insights on the assumptions of the kernel function in
@ref(eq:kernelcondition). The second one is that the proposed density
estimation method is only limited to the embedding points. Although for
the downstream task of anomaly detection, only density estimated on
these points is needed, it would be worth exploring the out-of-sample
density estimation and estimation on the boundary of manifolds
{[}@Berry2017-rz{]} in the future. The last question is related to the
quality of the embedding from dimension reduction methods. Although our
distortion correction method is fairly robust to different choices of
manifold learning methods, in certain cases, when the data structure is
too complex and the distortion is too severe to correct, the
quantitative relationship between embedding quality and density
estimation accuracy is not immediately clear. The embedding quality
could be measured using one of the metrics discussed in the online
supplementary material of @Cheng2021-dh, but when the ground truth
densities are unknown, which is usually the case with real-world data
set, it is hard to tell whether the distortion is corrected in the right
way. The density estimates on the edges of the whole data structure
could also be explored because most outer area points tend to be
detected as outliers. However, the outperformance of DC-KDE over KDE has
been shown in the higher dimensional simulation data and the electricity
usage data, which are more related to real-life data sets.

\hypertarget{acknowledgment}{%
\section*{Acknowledgment}\label{acknowledgment}}
\addcontentsline{toc}{section}{Acknowledgment}

This research was supported in part by the Monash eResearch Centre and
eSolutions-Research Support Services through the use of the MonARCH HPC
Cluster. The first author acknowledges the financial support of the
Monash Graduate Scholarship (MGS) and the Monash International Tuition
Scholarship (MITS) at Monash University.

\end{document}