### 2024-11-27 reports/20210603_fix_embedding.Rmd

---
title: "Riemannian metric from manifold learning"
author: "Fan Cheng"
date: "03/06/2021"
output: 
  # prettydoc::html_pretty:
    # theme: cayman
    # theme: flatly
    rmdformats::readthedown:
    toc: true
    float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, echo=TRUE, results='hide'}
library(tidyverse)
library(dimRed)
library(reticulate)
library(viridis)
library(hdrcde)
library(igraph)
library(matrixcalc)
library(akima)
library(car)
Jmisc::sourceAll(here::here("R/sources"))
set.seed(123)
```

# Smart meter data for one household

```{r data}
# load(here::here("data/spdemand_3639id336tow.rda"))
# nid <- 1
# ntow <- 336
# 
# train <- spdemand %>%
#   lazy_dt() %>%
#   filter(tow <= ntow,
#          # id <= sort(unique(spdemand[,id]))[nid],
#          id == 1003) %>%
#   dplyr::select(-id, -tow) %>%
#   data.table::as.data.table()
train <- readRDS(here::here("data/spdemand_1id336tow_train.rds"))
```

# Metric learning

## Radius searching with k-d trees

```{r parameters}
# Parameters fixed
x <- train
N <- nrow(train)
s <- 2
k <- 20
method <- "annIsomap"
annmethod <- "kdtree"
distance <- "euclidean"
treetype <- "kd"
searchtype <- "radius" # change searchtype for radius search based on `radius`, or KNN search based on `k`
radius <- .4 # the bandwidth parameter, \sqrt(\elsilon), as in algorithm
```

## The metric learning algorithm

```{r, message=FALSE}
metric_isomap <- metricML(x, s = s, k = k, radius = radius, method = method, annmethod = annmethod, eps = 0, distance = distance, treetype = treetype, searchtype = searchtype, invert.h = TRUE)
summary(metric_isomap)
```

The function `metricML()` returns a list of

-   the embedding coordinates `embedding` of $N \times s$
-   the embedding metric `rmetric` for each point $p \in x$ as an array
-   the weighted neighborhood graph as an `igraph` object `weighted_graph`
-   the sparse adjacency matrix from the graph `adj_matrix`
-   the graph laplacian matrix `laplacian`

## Check the output for each of the four steps

The metric learning algorithm consists of four steps:

1. construct a weighted neighborhood graph `weighted_graph`
2. calculate the graph laplacian `laplacian`
3. map the data $p \in x$ to the embedding coordinates `embedding`
4. apply the graph laplacian to the embedding corrdinates to obtain the embedding metric `rmetric`

Based on the output from megaman package, I compared the R function `metricML()` output for each of the four steps.

### Step1: weighted graph

We compare the graph by using the weighted adjacency matrix.

```{r, eval=T, include=T}
metric_isomap$adj_matrix[1:6, 1:6]
aff <-
  feather::read_feather(here::here("data/affinity_matrix_1id336tow.feather"))
aff[1:6, 1:6]
all.equal(as.matrix(aff), as.matrix(metric_isomap$adj_matrix), tolerance = 1e-5, check.attributes = F)
```

### Step2: Laplacian matrix

```{r, eval=T, include=T}
Ln <- metric_isomap$laplacian
Ln[1:6, 1:6]
lap <-
  feather::read_feather(here::here("data/laplacian_matrix_1id336tow.feather"))
lap[1:6, 1:6]
all.equal(as.matrix(lap), as.matrix(Ln), tolerance = 1e-5, check.attributes = F)
```

### Step3: Embedding coordinates

```{r, eval=T, include=T}
fn <- metric_isomap$embedding
fn %>% head()
emb_isomap <- feather::read_feather(here::here("data/embedding_isomap_1id336tow.feather"))
emb_isomap %>% head()
all.equal(as.matrix(emb_isomap), as.matrix(fn), tolerance = 1e-5, check.attributes = F)
par(mfrow = c(1, 2))
plot(fn, col = viridis::viridis(24), asp=1) 
plot(emb_isomap$`0`, emb_isomap$`1`, col = viridis::viridis(24), asp=1)
```

### Step4: Riemannian metric

In `metricML()`, the `invert.h` argument controls whether the inverse of Riemannian metric should be returned. FALSE by default.

The megaman package returns the inverse of Riemannian matrix.

```{r, eval=T, include=T}
# Rn <- riemann_metric(Y = fn, laplacian = Ln, d = s, invert.h = T)
Rn <- metric_isomap$rmetric
Rn[,,1:3]

np = import("numpy")
H_isomap <- np$load(here::here("data/hmatrix_isomap_1id336tow.npy"))
H_isomap[1,,, drop=TRUE] %>% 
  matrixcalc::is.positive.definite()
pyh_isomap <- array(NA, dim = c(s, s, N))
for(i in 1:N){
  pyh_isomap[,,i] <- H_isomap[i,,]
}
pyh_isomap[,,1:3] # inverted

all.equal(pyh_isomap, Rn, tolerance = 1e-5, check.attributes = F)

# If we use the embedding from Python, the inverse of Riem metric is the same. 
hn <- riemann_metric(Y = as.matrix(emb_isomap), laplacian = Ln, d = s, invert.h = T)
hn[,,1:3]
all.equal(pyh_isomap, hn, tolerance = 1e-5, check.attributes = F) # use python embedding
```
Finally, the R function `metricML()` gives the same/close outputs as the Python megaman package.

<!-- ## Embedding plots -->

<!-- ```{r} -->
<!-- # Comparison of Isomap embedding plot -->
<!-- par(mfrow = c(1, 2)) -->
<!-- plot(metric_isomap$embedding, col = viridis::viridis(24), asp=1) # metricML -->
<!-- emb_isomap <- -->
<!--   feather::read_feather(here::here("data/embedding_isomap_1id336tow.feather")) -->
<!-- plot(emb_isomap$`0`, emb_isomap$`1`, col = viridis::viridis(24), asp=1) # embedding from megaman, radius = 0.4 -->
<!-- ``` -->

<!-- ## Riemannian metric -->

<!-- ```{r} -->
<!-- # Use adjacency matrix as input for metricML and dimRed, then the embedding should be close, as well as the Riemannian metric -->
<!-- metric_isomap$adj_matrix[1:6, 1:6]# renormalized adjacency matrix -->
<!-- metric_isomap$weighted_graph %>% is.connected() -->
<!-- riem_isomap <- metric_isomap$rmetric -->
<!-- riem_isomap[,,1] %>%  -->
<!--   # isSymmetric() -->
<!--   matrixcalc::is.positive.definite() # FALSE -->
<!-- ``` -->

<!-- **The Riemannian metric from the `metricML()` function is now positive definite for all points.** -->

<!-- Now compare it with the megaman output. -->

<!-- ```{r, eval=TRUE} -->
<!-- np = import("numpy") -->
<!-- H_isomap <- np$load(here::here("data/hmatrix_isomap_1id336tow.npy")) -->
<!-- H_isomap[1,,, drop=TRUE] %>%  -->
<!--   matrixcalc::is.positive.definite() # TRUE -->

<!-- pd <- rep(NA, N) -->
<!-- for (i in 1:N) { -->
<!--   pd[i] <-  -->
<!--     riem_isomap[,,i] %>%  # R -->
<!--     # H_isomap[i,,, drop=TRUE] %>% # Python -->
<!--       # isSymmetric() -->
<!--       matrixcalc::is.positive.definite() # FALSE -->
<!-- } -->
<!-- pd %>% sum -->

<!-- # determinant(riem_isomap[,,i]) -->
<!-- # mat <- riem_isomap[,,i] -->
<!-- # eigen(mat) -->
<!-- ``` -->

# Ellipse plot

Using the Riemannian metric for each point as the 2-d covariance matrix and the 2-d embeddings as the centroid to plot an ellipse for each point.

The underlying requirement is that the Riemannian metric needs to be a square positive definite matrix, i.e. `matrixcalc::is.positive.definite()` returns `TRUE`.

```{r}
# TODO: if add = FALSE, fix xlim, ylim, radius for generating ellipse
# TODO: add argument `n.plot` to specify how many ellipse to be added
x <- metric_isomap$embedding
cols <- viridis::viridis(24) 
plot(metric_isomap$embedding, col = cols, asp=1, pch = 20)
for(i in 1:N){
  mat <- Rn[,,i] # pyh_isomap[,,I]
  center <- c(x[i,1], x[i,2])
  # add <- ifelse(i == 1, F, T)
  add <- T
  car::ellipse(center, mat, 
               radius = .1, # controls the size of ellipse, currently set manually
               # col = cols[i], 
               col = blues9[5],
               asp = 1, pty = "s", lwd = 1, center.pch = 19, center.cex = 0,
               fill = T, fill.alpha = 0.2, add = add, grid = T,
               # xlim = c(-.2, .25), ylim = c(-.2, .2)
               )
}
```

# Contour plots

## Variable kernel density estimate

For multivariate data, the variable kernel density estimate is given by $$\hat{f}(x) = n^{-1} \sum_i K_H (x - X_i).$$

### Inverse of Riemannian metric as bandwidth matrix

If we use the embedding `x` and the inverse of Riemannian metric `h`, we could produce contour plots based on the density estimates.

```{r pydensity, eval=T}
x <- cbind(emb_isomap$`0`, emb_isomap$`1`)
f <- mkde(x = x, h = pyh_isomap)
summary(f)
# Top 20 index with the lowest densities
head(order(f), n=20)
```

```{r, echo=FALSE, message=FALSE}
p_pyh <- plot.contour(x, f, plot.hdr = F)
```

Now apply the `mkde()` function to the output of `metricML()` written in R.

```{r rdensity}
x <- metric_isomap$embedding
riem_isomap <- metric_isomap$rmetric
f <- mkde(x = x, h = riem_isomap)
summary(f)
head(order(f), n=20)
```

```{r, echo=FALSE, message=FALSE}
p_rh <- plot.contour(x, f, plot.hdr = F)
```

We can check the outliers based on density `f`.
In R, the outliers are similar to the `megaman` output.

### Riemannian metric itself as bandwidth matrix

```{r pydensity1, eval=T}
R_isomap <- np$load(here::here("data/rmatrix_isomap_1id336tow.npy"))
# R_isomap[1,,, drop=TRUE] %>% 
#   matrixcalc::is.positive.definite() # not necessarily symmetric
pyr_isomap <- array(NA, dim = c(s, s, N))
for(i in 1:N){
  pyr_isomap[,,i] <- R_isomap[i,,]
}
pyr_isomap[,,1:3]

x <- cbind(emb_isomap$`0`, emb_isomap$`1`)
f <- mkde(x = x, h = pyr_isomap)
summary(f)
# Top 20 index with the lowest densities
head(order(f), n=20)
p_pyr <- plot.contour(x, f, plot.hdr = F)
```

Now apply the `mkde()` function to the output of `metricML()` written in R.

```{r rdensity1}
hn <- riemann_metric(Y = as.matrix(emb_isomap), laplacian = Ln, d = s, invert.h = T)
f <- mkde(x = metric_isomap$embedding, h = hn)
summary(f)
head(order(f), n=20)
p_rr <- plot.contour(x, f, plot.hdr = F)
```

# t-SNE

```{r tsne, message=FALSE}
metric_tsne <- metricML(x, s = s, k = k, radius = radius, method = "tSNE", annmethod = annmethod, eps = 0, distance = distance, treetype = treetype, searchtype = searchtype, invert.h = TRUE)
summary(metric_tsne)
```

```{r}
f <- mkde(x = metric_tsne$embedding, h = metric_tsne$rmetric)
summary(f)
head(order(f), n=20)
plot.contour(x = metric_tsne$embedding, f, plot.hdr = F)
```

```{r, echo=FALSE, eval=FALSE}
# Countour plot for megaman, dual metric
# x <- cbind(emb_isomap$`0`, emb_isomap$`1`)
embed_den <- as_tibble(cbind(x = x[,1], y = x[,2], z = f)) %>%
  drop_na()
pixel <- 100
lenbreak <- 5
akima.spl <- interp(embed_den$x, embed_den$y, embed_den$z, nx=pixel, ny=pixel, linear=FALSE)
hdrscatterplot(embed_den$x, embed_den$y, noutliers = 10)
filled.contour(akima.spl, color.palette = viridis,
               plot.axes = { axis(1); axis(2);
                 title("smooth interp(*, linear = FALSE)");
                 points(embed_den, pch = 3, col= hcl(c=20, l = 10))})
```

```{r, echo=FALSE, include=FALSE, eval=FALSE}
embed_den <- as_tibble(cbind(x = x[,1], y = x[,2], z = f)) %>%
  drop_na()
# library(akima)
pixel <- 100
lenbreak <- 5
akima.spl <- interp(embed_den$x, embed_den$y, embed_den$z, nx=pixel, ny=pixel, linear=FALSE)
p.zmin <- min(embed_den$z,na.rm=TRUE)
p.zmax <- max(embed_den$z,na.rm=TRUE)
breaks <- pretty(c(p.zmin,p.zmax), lenbreak)
# hdrcde
hdrscatterplot(embed_den$x, embed_den$y, noutliers = 10)
hdrinfo <- hdr.2d(embed_den$x, embed_den$y, prob = c(50, 95, 99))
plot.hdr2d(hdrinfo)
# Contour
contour(akima.spl, main = "smooth interp(*, linear = FALSE)", col = viridis(length(breaks)-1), levels=breaks, add=TRUE)
points(embed_den, pch = 20)

filled.contour(akima.spl, color.palette = viridis,
               plot.axes = { axis(1); axis(2);
                 title("smooth interp(*, linear = FALSE)");
                 points(embed_den, pch = 3, col= hcl(c=20, l = 10))},
               # plot.title = {par(cex.main=1);title(main = "The Topography of Maunga Whau", xlab = "Meters North", ylab = "Meters West")},
               # plot.axes = { axis(1, seq(100, 800, by = 100))
                 # axis(2, seq(100, 600, by = 100)) },
               # key.title = {par(cex.main=0.3);title(main="Height\n(meters)")},
               # key.axes = axis(4, seq(-1, 1.5, by = 0.5), asp=1)
)

```