### 2024-11-27 reports/20210617_plots.Rmd

---
title: "Plotting for metric learning output"
author: "Fan Cheng"
date: "2021-06-17"
output: 
  # prettydoc::html_pretty:
    # theme: cayman
    # theme: flatly
    rmdformats::readthedown:
    toc: true
    float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, echo=TRUE, results='hide'}
library(tidyverse)
library(dimRed)
library(reticulate)
library(here)
library(viridis)
library(hdrcde)
library(igraph)
library(matrixcalc)
library(akima)
library(car)
library(ggforce)
library(ks)
Jmisc::sourceAll(here::here("R/sources"))
set.seed(123)
```

# Metric learning

## Smart meter data for one household

```{r data}
# load(here::here("data/spdemand_3639id336tow.rda"))
# nid <- 1
# ntow <- 336
# 
# train <- spdemand %>%
#   lazy_dt() %>%
#   filter(tow <= ntow,
#          # id <= sort(unique(spdemand[,id]))[nid],
#          id == 1003) %>%
#   dplyr::select(-id, -tow) %>%
#   data.table::as.data.table()
train <- readRDS(here::here("data/spdemand_1id336tow_train.rds"))
```

## Radius searching with k-d trees

```{r parameters}
# Parameters fixed
x <- train
N <- nrow(train)
s <- 2
k <- 20
method <- "annIsomap"
annmethod <- "kdtree"
distance <- "euclidean"
treetype <- "kd"
searchtype <- "radius" # change searchtype for radius search based on `radius`, or KNN search based on `k`
radius <- .4 # the bandwidth parameter, \sqrt(\elsilon), as in algorithm
```

## The metric learning algorithm

```{r, message=FALSE}
metric_isomap <- metricML(x, s = s, k = k, radius = radius, method = method, annmethod = annmethod, eps = 0, distance = distance, treetype = treetype, searchtype = searchtype, invert.h = TRUE)
summary(metric_isomap)
```

The function `metricML()` returns a list of

-   the embedding coordinates `embedding` of $N \times s$
-   the embedding metric `rmetric` for each point $p \in x$ as an array. By default, `invert.h = TRUE` returns the inverse of Riemannian metric. 
-   the weighted neighborhood graph as an `igraph` object `weighted_graph`
-   the sparse adjacency matrix from the graph `adj_matrix`
-   the graph laplacian matrix `laplacian`

# Embedding plot

```{r embedplot, eval=FALSE, include=FALSE}
# Embedding plot of metricML() output for electricity data
plot_embedding <- function(x) {
  fn <- x$embedding

  todcolor <- colorspace::scale_color_continuous_sequential(
    palette = "viridis",
    breaks = c(0, 12, 24, 36, 48),
    labels=c("00:00", "06:00", "12:00", "18:00", "24:00"),
    name="Time of day",
    guide=guide_colorbar(barwidth = 10))

  p <- fn %>% 
    cbind(tod = rep(1:48, N / 48)) %>% 
    as_tibble() %>% 
    ggplot(aes(x = E1, y = E2, col = tod)) + 
    geom_point() + 
    # labs(x = "ISO1", y = "ISO2") +
    coord_fixed(ratio = 1) + 
    todcolor + 
    theme(legend.position = 'bottom')

  return(p)
}
```

```{r, eval=T, include=T}
# fn <- metric_isomap$embedding
# head(fn)
# plot(fn, col = viridis::viridis(24), asp = 1)
plot_embedding(metric_isomap) +
  labs(x = "ISO1", y = "ISO2")
```

# Ellipse plot

Using the Riemannian metric for each point as the 2-d covariance matrix and the 2-d embeddings as the centroid to plot an ellipse for each point.

The underlying requirement is that the Riemannian metric needs to be a square positive definite matrix.

```{r car, include=FALSE, eval=FALSE}
# Done: add argument `n.plot` to specify how many ellipse to be added
# If add = FALSE, fix xlim, ylim, radius for generating ellipse
plot.ellipse <- function(x, n.plot = 50) {
  fn <- x$embedding
  Rn <- x$rmetric
  cols <- viridis::viridis(24) 
  n.plot <- 50
  plot(fn, col = cols, asp = 1, pch = 20)
  for (i in sample(1:N, n.plot)) {
    mat <- Rn[,,i] 
    center <- c(fn[i,1], fn[i,2])
    add <- T 
    car::ellipse(center, mat, 
                 radius = .1, # controls the size of ellipse, currently set manually
                 # xlim = c(-.2, .25), ylim = c(-.2, .2),
                 # col = cols[i],
                 col = blues9[5],
                 asp = 1, pty = "s", lwd = 1, center.pch = 19, center.cex = 0,
                 fill = T, fill.alpha = 0.2, grid = T, add = add, 
                 )
  }
}
plot.ellipse(metric_isomap)
```

```{r ggellipse}
plot_ellipse(metric_isomap, add = F, n.plot = 50, scale = 20, 
             color = blues9[5], fill = blues9[5], alpha = 0.2)
# fn <- metric_isomap$embedding
# Rn <- metric_isomap$rmetric # array
# n.plot <- 50
# e <- riem2ellipse(Rn, scale = 20) %>% 
#   cbind(fn) %>% 
#   as_tibble()
# OR
# plot_embedding(metric_isomap, embedding = F) +
#   plot_ellipse(metric_isomap, add = T)
# OR
# # plot_embedding(fn, embedding = T) + 
#   ggplot(data = e, aes(E1, E2)) +
#   geom_point() +
#   geom_ellipse(data = slice_sample(e, n = n.plot), 
#                aes(x0 = E1, y0 = E2, a = a, b = b, angle = angle, group = -1), 
#                color = blues9[5], fill = blues9[5], alpha = 0.2)
```

# Variable kernel density estimate

For multivariate data, the variable kernel density estimate is given by $$\hat{f}(x) = n^{-1} \sum_i K_{H_i} (x - X_i).$$

## Outlier plot based on density estimates

### Fixed bandwidth

```{r}
# fixed bandwidth
fn <- metric_isomap$embedding
E1 <- fn[,1] # rename as Ed to match the aesthetics in plot_ellipse()
E2 <- fn[,2]
prob <- c(1, 50, 99)
p_outlier <- hdrscatterplot(E1, E2, levels = prob, noutliers = 20, label = NULL) + 
  plot_ellipse(metric_isomap, add = T, n.plot = 50, scale = 20, 
             color = blues9[5], fill = blues9[5], alpha = 0.2)
p_outlier
```

### Pointwise adaptive bandwidth

```{r outliers}
Rn <- metric_isomap$rmetric # array
n.grid <- 10
f <- vkde2d(x = fn[,1], y = fn[,2], h = Rn, n = n.grid)
plot_contour(metric_isomap, n.grid = n.grid, scale = 0.1)
```

```{r hdroutliers}
source(here::here("R/sources/hdrplotting.R"))
p_outlier_vkde <- plot_outlier(x = metric_isomap, n.grid = 50, prob = prob)
# den <- f
# x <- E1; y <- E2
#  <-  = c(50, 95, 99)
# # prob = (1:10)*10
# # Convert prob to coverage percentage if necessary
# if(max(prob) > 50) {# Assume prob is coverage percentage
#   alpha <- (100-prob)/100
# } else {# prob is tail probability (for backward compatibility)
#   alpha <- prob}
# alpha <- sort(alpha)
# # Calculates falpha needed to compute HDR of bivariate density den.
# # Also finds approximate mode.
# fxy <- hdrcde:::interp.2d(den$x,den$y,den$z,x,y)
# falpha <- quantile(fxy, alpha)
# index <- which.max(fxy)
# mode <- c(x[index],y[index])
# hdr2d_info <- structure(list(mode=mode,falpha=falpha,fxy=fxy, den=den, alpha=alpha, x=x, y=y), class="hdr2d") # list for hdr.2d() output
# # plot.hdr2d(hdr2d_info, show.points = T, outside.points = T, pointcol = grey(0.5), xlim = round(range(x)), ylim = round(range(y)))
# 
# p_outlier_vkde <- hdrscatterplot(E1, E2, levels = prob, noutliers = 20, label = NULL, den = hdr2d_info)  + 
#   plot_ellipse(metric_isomap, add = T, n.plot = 50, scale = 20, 
#              color = blues9[5], fill = blues9[5], alpha = 0.2)
```

```{r compoutlier}
library(patchwork)
(p_outlier + p_outlier_vkde ) + coord_fixed()
```

```{r, include=FALSE}
# todcolor <- colorspace::scale_color_continuous_sequential(
#   palette = "viridis",
#   # breaks = c(0, 12, 24, 36, 48),
#   # labels=c("00:00", "06:00", "12:00", "18:00", "24:00"),
#   name="Density",
#   guide=guide_colorbar(barwidth = 10))
# 
# p <- fn %>% 
#   # cbind(tod = rep(1:48, N / 48)) %>%
#   as_tibble() %>% 
#   ggplot(aes(x = E1, y = E2)) + 
#   geom_point() + 
#   # labs(x = "ISO1", y = "ISO2") +
#   coord_fixed(ratio = 1) + 
#   # todcolor + 
#   theme(legend.position = 'bottom')
#   
# p + 
#   plot_ellipse(metric_isomap, add = T, n.plot = 50, scale = 20, 
#              color = blues9[5], fill = blues9[5], alpha = 0.2)
```

```{r, include=FALSE, eval=FALSE}
fn <- metric_isomap$embedding
hdrscatterplot(fn[,1], fn[,2], noutliers = 10)
hdrinfo <- hdr.2d(fn[,1], fn[,2], prob = prob)
plot.hdr2d(hdrinfo)

# f <- vkde2d(x = fn[,1], y = fn[,2], h = Rn, n = 20)
# # modified
# f1 <- den.estimate.2d(x = fn[,1], y = fn[,2], kde.package = "ks", h = Rn[,,1], xextend =0.15, yextend=0.15)  # only one bandwidth at a time
# hdrinfo_h <- hdr.2d(fn[,1], fn[,2], prob = (1:20)*5, den = f1)
# plot.hdr2d(hdrinfo_h)

Rh <- apply(Rn, 3, det)
vfhat.sp <- ks:::kde.sp.2d(x=fn, H=Rh)
plot(vfhat.sp, display="persp", box=TRUE, phi=20, zlim=c(0,max(vfhat.sp$estimate)))

# den <- den.estimate.2d(x=fn[,1], y=fn[,2], kde.package="ks", h=Rh)
x=fn[,1]; y=fn[,2]
den <- vfhat.sp
den <- list(x=den$eval.points[[1]],y=den$eval.points[[2]],z=den$estimate)

prob = c(50, 95, 99)
prob = (1:10)*10
# Convert prob to coverage percentage if necessary
if(max(prob) > 50) {# Assume prob is coverage percentage
  alpha <- (100-prob)/100
} else {# prob is tail probability (for backward compatibility)
  alpha <- prob}
alpha <- sort(alpha)
# Calculates falpha needed to compute HDR of bivariate density den.
# Also finds approximate mode.
fxy <- hdrcde:::interp.2d(den$x,den$y,den$z,x,y)
falpha <- quantile(fxy, alpha)
index <- which.max(fxy)
mode <- c(x[index],y[index])
hdr2d_info <- structure(list(mode=mode,falpha=falpha,fxy=fxy, den=den, alpha=alpha, x=x, y=y), class="hdr2d") # list for hdr.2d() output
plot.hdr2d(hdr2d_info, show.points = F, outside.points = T, pointcol = grey(0.5), xlim = round(range(x)), ylim = round(range(y)))

filled.contour(den, color.palette = viridis,
                      plot.axes = { axis(1); axis(2);
                        points(fn, pch = 3, col= hcl(c=20, l = 8))})
```

```{r, eval=FALSE, include=FALSE}
fn <- metric_isomap$embedding
Rn <- metric_isomap$rmetric
# Variable KDE
data(worldbank)
wb <- as.matrix(na.omit(worldbank[,4:5]))
xmin <- c(-70,-25); xmax <- c(35,70)
fhat <- kde(x=wb, xmin=xmin, xmax=xmax)
fhat.sp <- kde.sp(x=wb, xmin=xmin, xmax=xmax)
vfhat.sp <- ks:::kde.sp.2d(x=wb, xmin=xmin, xmax=xmax, H = fhat.sp$H)
plot(fhat, display="persp", box=TRUE, phi=20, zlim=c(0,max(fhat$estimate)))
plot(fhat.sp, display="persp", box=TRUE, phi=20, zlim=c(0,max(fhat.sp$estimate)))
plot(vfhat.sp, display="persp", box=TRUE, phi=20, zlim=c(0,max(vfhat.sp$estimate)))
## Not run:
fhat.ball <- kde.balloon(x=wb, xmin=xmin, xmax=xmax)
plot(fhat.ball, display="persp", box=TRUE, phi=20, zlim=c(0,max(fhat.sp$estimate)))

ks::Hpi.diag(fn,binned=TRUE)
Hns(fn, deriv.order=4)

fhat.sp$H # vector of grid point length, not matrices
```

<!-- # Contour plot -->

<!-- ### Inverse of Riemannian metric as bandwidth matrix -->

<!-- If we use the embedding `x` and the inverse of Riemannian metric `h`, we could produce contour plots based on the density estimates. -->

<!-- ```{r pydensity, eval=F, include=FALSE} -->
<!-- emb_isomap <- feather::read_feather(here::here("data/embedding_isomap_1id336tow.feather")) -->
<!-- np <- reticulate::import("numpy") -->
<!-- H_isomap <- np$load(here::here("data/hmatrix_isomap_1id336tow.npy")) -->
<!-- pyh_isomap <- array(NA, dim = c(s, s, N)) -->
<!-- for(i in 1:N){ -->
<!--   pyh_isomap[,,i] <- H_isomap[i,,] -->
<!-- } -->
<!-- x <- cbind(emb_isomap$`0`, emb_isomap$`1`) -->
<!-- # f <- mkde(x = x, h = pyh_isomap) -->
<!-- # Top 20 index with the lowest densities -->
<!-- # head(order(f), n = 20) -->
<!-- plot_contour(list(embedding = x, rmetric = pyh_isomap), n.grid = 10) -->
<!-- ``` -->

<!-- ```{r contour} -->
<!-- plot_contour(metric_isomap, n.grid = 20) -->
<!-- # fn <- metric_isomap$embedding -->
<!-- # Rn <- metric_isomap$rmetric # array -->
<!-- # #h <- t(apply(Rn, 3, diag)) -->
<!-- # f <- vkde2d(x = fn[,1], y = fn[,2], h = Rn, n = 100) -->
<!-- # str(f) -->
<!-- # # image(f) -->
<!-- # p <- filled.contour(f, color.palette = viridis, -->
<!-- #                  plot.axes = { axis(1); axis(2); -->
<!-- #                    points(fn, pch = 3, col= hcl(c=20, l = 8))}) -->
<!-- ``` -->

<!-- ### Riemannian metric itself as bandwidth matrix -->

<!-- ```{r rdensity1} -->
<!-- fn <- metric_isomap$embedding -->
<!-- s <- 2 -->
<!-- hn <- riemann_metric(Y = fn, laplacian = metric_isomap$laplacian, d = s, invert.h = F) -->
<!-- # f <- vkde2d(x = fn[,1], y = fn[,2], h = hn, n = 20) -->
<!-- # str(f) -->
<!-- # # image(f) -->
<!-- # p <- filled.contour(f, color.palette = viridis, -->
<!-- #                  plot.axes = { axis(1); axis(2); -->
<!-- #                    points(fn, pch = 3, col= hcl(c=20, l = 8))}) -->
<!-- plot_contour(list(embedding = fn, rmetric = hn), n.grid = 100) -->
<!-- ``` -->

# t-SNE

```{r tsne, message=FALSE, warning=FALSE, eval=TRUE}
x <- train
method <- "tSNE"
perplexity <- 30
theta <- 0.5
metric_tsne <- metricML(x, s = s, k = k, radius = radius, method = method, annmethod = annmethod, eps = 0, distance = distance, treetype = treetype, searchtype = searchtype, perplexity = perplexity, theta = theta, invert.h = TRUE)
summary(metric_tsne)
```

```{r, message=FALSE, eval=TRUE}
ftsne <- metric_tsne$embedding
E1 <- ftsne[,1]; E2 <- ftsne[,2]
# plot_embedding(metric_tsne)
plot_ellipse(metric_tsne, n.plot = 50)
plot_contour(metric_tsne, n.grid = 20, scale = 1/8)
p_tsne <- plot_outlier(x = metric_tsne, n.grid = 20, prob = prob, noutliers = 20, scale = 1/8)
p_tsne
hdrscatterplot(E1, E2, kde.package = "ks", noutliers = 20) + 
  plot_ellipse(metric_tsne, n.plot = 50, add = T)
p_outlier_vkde + p_tsne
```

```{r, include=FALSE, eval=FALSE}
perplexity <- 30
theta <- 0.5
method <- "tSNE"
e <- dimRed::embed(x,
                     .method = method,
                     knn = k,
                     ndim = s,
                     annmethod = annmethod,
                     radius = radius,
                     eps = eps,
                     nt = nt,
                     nlinks = nlinks,
                     ef.construction = ef.construction,
                     distance = distance,
                     treetype = treetype,
                     searchtype = searchtype,
                     perplexity = perplexity,
                     theta = theta,
                     .mute = c("output"),
                     # ...

  )
  fn <- e@data@data
plot(metric_tsne$embedding)
```

```{r anntsne, message=FALSE, warning=FALSE, eval=FALSE, echo=FALSE}
x <- train
nn2res <- dplyr::case_when(distance=="euclidean" ~ 
                                 RANN::nn2(data = x, query = x, k = k + 1, 
                                           treetype = treetype, searchtype = searchtype, eps = eps,
                                           radius = radius),
                               distance=="manhattan" ~ 
                                 RANN.L1::nn2(data = x, query = x, k = k + 1, 
                                              treetype = treetype, searchtype = searchtype, eps = eps,
                                              radius = radius),
    )
    names(nn2res) <- c("nn.idx", "nn.dists")

metric_tsne1 <- metricML(x, nn.idx = nn2res$nn.idx, nn.dists = nn2res$nn.dists, s = s, k = k, radius = radius, method = "anntSNE", annmethod = annmethod, eps = 0, distance = distance, treetype = treetype, searchtype = searchtype, perplexity = k-1, theta = 0.5, invert.h = TRUE)
summary(metric_tsne1)

plot.contour(metric_tsne1, n.grid = 10)
```

```{r, echo=FALSE, include=FALSE, eval=FALSE}
embed_den <- as_tibble(cbind(x = x[,1], y = x[,2], z = f)) %>%
  drop_na()
# library(akima)
pixel <- 100
lenbreak <- 5
akima.spl <- interp(embed_den$x, embed_den$y, embed_den$z, nx=pixel, ny=pixel, linear=FALSE)
p.zmin <- min(embed_den$z,na.rm=TRUE)
p.zmax <- max(embed_den$z,na.rm=TRUE)
breaks <- pretty(c(p.zmin,p.zmax), lenbreak)
# hdrcde
hdrscatterplot(embed_den$x, embed_den$y, noutliers = 10)
hdrinfo <- hdr.2d(embed_den$x, embed_den$y, prob = c(50, 95, 99))
plot.hdr2d(hdrinfo)
# Contour
contour(akima.spl, main = "smooth interp(*, linear = FALSE)", col = viridis(length(breaks)-1), levels=breaks, add=TRUE)
points(embed_den, pch = 20)

filled.contour(akima.spl, color.palette = viridis,
               plot.axes = { axis(1); axis(2);
                 title("smooth interp(*, linear = FALSE)");
                 points(embed_den, pch = 3, col= hcl(c=20, l = 10))},
               # plot.title = {par(cex.main=1);title(main = "The Topography of Maunga Whau", xlab = "Meters North", ylab = "Meters West")},
               # plot.axes = { axis(1, seq(100, 800, by = 100))
                 # axis(2, seq(100, 600, by = 100)) },
               # key.title = {par(cex.main=0.3);title(main="Height\n(meters)")},
               # key.axes = axis(4, seq(-1, 1.5, by = 0.5), asp=1)
)

```