### 2024-11-27 reports/20210729_review.Rmd

---
title: "July Summary"
author: "Fan Cheng"
date: "2021-07-28"
output: 
  # prettydoc::html_pretty:
    # theme: cayman
    # theme: flatly
    rmdformats::readthedown:
    toc: true
    float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Supervision form

- External supervisor form needed

- Second Monash supervisor

# July summary

Below is a summary of some events I attended with some feedback.

## ANZSC 2021

Machine learning session; 10min talk + 5min Q&A.

- How to define anomalies?

- Anomalies sensitivity to manifold learning methods.

- Examine distributions of nearby typical instead of anomalous households.

## AMSI winter school

### Lectures

- More insights about dimension reduction from Tas.

- A lot of Bayesian statistics.

### Runner-up for the participants talks

- Two rounds: 8 people group, 13 finalist; vote for top 3 talks; 
- 10min talk, no live questions;

- skipped manifold learning and quality measures after the lectures

### Possible collaborations

ABS: Victoria smart meter data project, detailed to be discussed.

> Brian Chan: Ohh interesting, is the Vic data publicly available? The reason why I mention it is because I know we are doing some very interesting work with Smart Meter Data (not my team so I don't know the details) - perhaps there might be some collaboration opportunities here, I can definitely follow-up if you are interested

> Owen Forbes: how I might use manifold learning with my EEG/neuroscience data for my PhD. 
One approach I am considering for my future work is representing each EEG frequency spectrogram as a curve - could potentially use functional PCA where each channel’s spectrogram is a curve. But seeing your application to electrical usage patterns, it seems comparable not necessarily to the EEG time series but to the power as a function of frequency for each electrode

<!-- , kind of like this https://www.google.com/url?sa=i&url=https%3A%2F%2Fwww.researchgate.net%2Ffigure%2FPS[…]e=images&cd=vfe&ved=0CAsQjRxqFwoTCODswZzE7vECFQAAAAAdAAAAABAF -->

## ECSSC 2021

A 20min talk scheduled at Friday 11:20 am.

More time on manifold learning, quality measures and the result plots.

## MLANN paper revision

- Added t-SNE and UMAP implementation.

- Introduction: add references for t-SNE and UMAP; more explanations on why embedding quality matters more than recall rate; add references on t-SNE variations and UMAP using NN-Descent for fast NN and their limitation [?].

- Manifold learning: explain parameter settings for $K$ and $d$; t-SNE and UMAP; UMAP algorithm pointed to the paper [?].

- Quality measures: recall rate definition and why high recall rate is not enough for preserving accurate local structure [any insights?].

- Unify embedding dimension as $d=2$ for a better-visualization reason. Plus, t-SNE only works for $d=1,2,3$ since it's designed mainly for high-dimensional data visualization with scatterplots.

- MNIST experiments: list ML parameter settings; Update $d=2$ result.

- Smart meter data application: clarify dimension $p=200$ and N for different scenarios.

- Conclusions: extension is possible for other methods (ML & ANN) to avoid adding other methods later; add intrinsic dimension reference from Chris [?].